---
title: Final Report
toc: true
toc-title: Table of contents
toc-location: left
params:
  number: 3
  weight: '16%'
  purpose: 'purpose'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  fig.path = "figs/",
  fig.width = 7.252,
  fig.height = 4,
  comment = "#>",
  fig.retina = 3
)

source(here::here("R", "settings.R"))

# Read in rubric data
rubric <- read_csv(here::here('rubrics', 'final-report.csv'))
maxPoints <- rubric %>%
    filter(rating == "Excellent") %>%
    summarise(max = sum(maxPoints)) %>%
    pull(max)
```

```{r child = here::here("fragments", "project.qmd")}
```

Your final report should be a fully reproducible html webpage. It should include text, data, code, and plots. Your report should be written in a narrative format (i.e. using coherent paragraphs rather than a series of bullet points). You may use headings where appropriate to break up your report into sections. There is no length requirement - your report should be sufficiently detailed to address the requirements listed below and sufficiently concise such that it is expressed in the fewest necessary words. It is also perfectly acceptable for large portions of your final report to be identical to sections in your progress report, albeit all results should be "polished" and publication ready. Below is a list of specific items your report should include.

## 1. Get organized

1. Download and unzip [this template](../templates/project-template.zip) for your proposal report. Open the `project.Rproj` file and write your progress report in the `report.Rmd` file. The template comes with some text and code explaining how to use it - **you should delete the template code and text as it is only for explanatory purposes**. Be sure to adjust the content in the YAML:

    - Write your project title in the `title` field (and provide a subtitle if you wish, or delete the `subtitle` field).
    - In the `author` field, list the names of all teammates, e.g. `author: Luke Skywalker, Leia Organa, and Han Solo`.

2. Put the data you are working with for your project in the appropriate folders. Even if you are reading in data directly from a website (e.g. `data <- read_csv("www.somewhere.com/data.csv)`), be sure to download a local copy of the data and put it in the appropriate data folder.

3. At the bottom of your report, add an Appendix section that includes data dictionaries for the key data sets used in your analysis as well as a copy of all of the code used in your report, which can be automatically generated with a single chunk like this:

````markdown
`r ''````{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
````

## 2. State your research question

Clearly state your research question and motivate why it is important / why the reader should care about it. You should follow [these guidelines](https://writingcenter.gmu.edu/guides/how-to-write-a-research-question).

## 3. Discuss your data sources

Discuss the data source(s) you are using for your analysis. For each data source:

- Describe it and include urls or references to the original sources as well as links to any pre-processed or formatted data you are using. These are not always the same. For example, in the [Mini Project 1 assignment](a-mini-project-1.html), the original source was the [Transit Costs Project](https://transitcosts.com/) whereas the data file used was posted on [this GitHub repo](https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-01-05).
- Discuss the validity of your data:

    - Is the data you are using from the _original_ source, or has it been pre-processed?
    - How was the original data collected and by whom?
    - Are there perhaps any missing data that might not have been observed?
    - Could the data be biased in some way?

- Finally, provide a data dictionary for each data file you are using in an appendix at the end of your report. The dictionary should contain a table of each variable name along with a brief description of the variable.

## 4. Describe your results with text and charts

- Display polished charts that either support or oppose your research question or illustrate what else you might need to address your research question.
- Your charts should be "polished" with appropriate labels and follow the design principles we have covered in class.
- Your plot type choices should highlight the main point(s) you want to make or clearly show the relationship you want to emphasize.
- Around your charts, write narrative text explaining what the chart shows and its significance towards addressing your research question. This should read as a continuous story rather than as a reply to each of the requirements described here.
- Your analysis must include at **least three polished charts / animations**, though you can of course include more. Alternatively, you can build an interactive shiny app with at least two dynamic charts. If you build a shiny app, include a link to your app and screenshots of key charts from your app in your final report.

## 5. Review your teammates

(this is one of the only things I'll ask for on Blackboard)

On Blackboard under the assignment titled "Team Member Review: Final Report", write a short description of the specific contributions the team member made throughout the whole project (from proposal to the final report). Note down if there are any team members you feel have done an exceptional job (gone beyond what was required) or have not contributed in a significant way (e.g. does not respond to team communications, turns in parts they are responsible for late or never).

These reviews will be kept confidential and compared to assess that the workload and grading for team members are equitable. Team members who do not make meaningful contributions to their projects will not receive the same grade as that of their teammates.

Here is an example review:

```
John:

- Identified the data source and wrote the documentation for it.
- Helped with initial data cleaning.you
- Wrote introduction section of all reports.

Jane:

 - Completed final data cleaning process.
 - Did most of the initial data exploration, including most of the charts in the progress report.

Paul:

- Wrote code for the final, polished visualizations.
- Did most of the remaining writing of the final report.
```

## 6. Knit and submit

1. Click the "knit" button to compile your `.Rmd` file into a html web page.
2. Spell check your report by running: `spelling::spell_check_files("report.Rmd")`.
3. Proofread your html webpage! If any fixes are needed, repeat steps 1 & 2.
4. Create a zip file of everything in your R Project folder. Name your file `final-report.zip`, then go to your team Box folder and submit your zip file in the "submissions" folder. **Only one person from your team should submit the report.**

# Grading Rubric

### `r maxPoints` Total Points

```{r, echo=FALSE}
rubric %>%
  mutate(description = paste0("<b>", points, '</b><br>', description)) %>%
  select(-points) %>%
  spread(key = rating, value = description) %>%
  select(-category) %>%
  rename(Category = label) %>%
  arrange(order) %>%
  select(-order) %>%
  select(-maxPoints) %>%
  kable(format = 'html', escape = FALSE) %>%
  kable_styling(bootstrap_options = "striped")
```

