---
title: "Amounts & Proportions"
subtitle: "EMSE 4572: Exploratory Data Analysis"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# install.packages("waffle")
library(tidyverse)
library(here)
library(cowplot)
library(waffle)
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3
)

# Read in data
fed_spending_wide <- read_csv(here('data', 'fed_spending_wide.csv'))
fed_spending_long <- read_csv(here('data', 'fed_spending_long.csv'))
milk_prod_wide <- read_csv(here('data', 'milk_prod_wide.csv'))
milk_prod_long <- read_csv(here('data', 'milk_prod_long.csv'))
avengers         <- read_csv(here('data', 'avengers.csv'))
bears            <- read_csv(here('data', 'north_america_bear_killings.csv'))
federal_spending <- read_csv(here('data', 'fed_spending_long.csv'))
gapminder        <- read_csv(here('data', 'gapminder.csv'))
lotr_words       <- read_csv(here('data', 'lotr_words.csv'))
milk_production  <- read_csv(here('data', 'milk_production.csv'))
wildlife_impacts <- read_csv(here('data', 'wildlife_impacts.csv'))
```

# Tidy data review

```{r}
# Tidy the data
```

```{r}
# Visualize variability
```

# Practice manipulating factors

Use the `wildlife_impacts` data to create the following plot

```{r wildlife_phase_of_flight_bars, fig.width=9, fig.height=5}
# Write code here
```

## Practice plotting amounts

Create the following charts:

```{r bear_bars, fig.width=7, fig.height=5}
# Write code here
```

```{r milk_dots_top10, fig.height=4, fig.width=7}
# Write code here
```

## Your turn - practice plotting proportions

Using the `wildlife_impacts` data, create plots that shows the proportion of incidents that occur at each different time of day.

For this exercise, you can remove `NA` values.

Try to create the following plots:

- Stacked bars
- Dodged bars
- Pie chart
- Waffle chart

```{r}
wildlife_summary <- wildlife_impacts %>%
  filter(!is.na(time_of_day)) %>%
  count(time_of_day)

wildlife_summary
```

Stacked bars:

```{r wildlife_bars_stacked, fig.width=6, fig.height=3}
# Write code here
```

Dodged bars:

```{r wildlife_bars_dodged, fig.width=5, fig.height=3}
# Write code here
```

Pie chart:

```{r wildlife_pie, fig.width=5, fig.height=3}
# Write code here
```

Waffle chart:

```{r wildlife_waffle, fig.height=5, fig.width=3.5}
# Write code here
```
