---
title: "Amounts & Proportions"
subtitle: "EMSE 4572: Exploratory Data Analysis"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# install.packages("waffle")
library(tidyverse)
library(here)
library(cowplot)
library(waffle)
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3
)

# Read in data
fed_spending_wide <- read_csv(here('data', 'fed_spending_wide.csv'))
fed_spending_long <- read_csv(here('data', 'fed_spending_long.csv'))
milk_prod_wide <- read_csv(here('data', 'milk_prod_wide.csv'))
milk_prod_long <- read_csv(here('data', 'milk_prod_long.csv'))
avengers         <- read_csv(here('data', 'avengers.csv'))
bears            <- read_csv(here('data', 'north_america_bear_killings.csv'))
federal_spending <- read_csv(here('data', 'fed_spending_long.csv'))
gapminder        <- read_csv(here('data', 'gapminder.csv'))
lotr_words       <- read_csv(here('data', 'lotr_words.csv'))
milk_production  <- read_csv(here('data', 'milk_production.csv'))
wildlife_impacts <- read_csv(here('data', 'wildlife_impacts.csv'))
```

# Tidy data review

```{r}
# Tidy the data
fed_spending_long <- fed_spending_wide %>%
    gather(key = "department", value = "rd_budget_mil", DHS:VA)

milk_prod_long <- milk_prod_wide %>%
    gather(key = "region", value = "milk_production", Appalachian:Pacific)
```

```{r}
# Visualize variability
fed_spending_long %>%
    ggplot() +
    geom_histogram(aes(x = rd_budget_mil))

milk_prod_long %>%
    ggplot() +
    geom_histogram(aes(x = milk_production))
```

# Practice manipulating factors

Use the `wildlife_impacts` data to create the following plot

```{r wildlife_phase_of_flight_bars, fig.width=9, fig.height=5}
wildlife_impacts %>%
  filter(!is.na(phase_of_flt)) %>%
  mutate(phase_of_flt = str_to_title(phase_of_flt)) %>%
  count(phase_of_flt) %>%
  mutate(
    phase_of_flt = fct_other(phase_of_flt, keep = c(
      "Approach", "Take-Off Run", "Landing Roll", "Climb")),
    phase_of_flt = fct_reorder(phase_of_flt, n),
    phase_of_flt = fct_recode(phase_of_flt, "All Other" = "Other")) %>%
  ggplot() +
  geom_col(aes(x = n, y = phase_of_flt), width = 0.7) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +
  theme_minimal_vgrid(font_size = 18) +
  labs(x = "Count",
       y = "Phase of flight",
       title = 'Number of wildlife impacts by phase of flight')
```

## Your turn - practice plotting amounts

Create the following charts:

```{r bear_bars, fig.width=7, fig.height=5}
bears %>%
  count(month) %>%
  mutate(
    month = as.factor(month),
    is_aug = ifelse(month == "8", TRUE, FALSE)) %>%
  ggplot() +
  geom_col(
    aes(x = month, 
        y = n,
        fill = is_aug),
    width = 0.7) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    limits = c(0, 30)) +
  scale_fill_manual(values = c("gray", "red")) +
  theme_minimal_hgrid(font_size = 18) +
  theme(legend.position = "none") +
  labs(
      x = "Month",
      y = "Count",
      title = 'Number of bear killings in each month',
      subtitle = 'August is the deadliest month'
  )
```

```{r milk_dots_top10, fig.height=4, fig.width=7}
milk_production %>%
  filter(year == 2017) %>%
  arrange(desc(milk_produced)) %>%
  slice(1:10) %>%
  mutate(
    milk_produced = milk_produced / 10^9,
    state = fct_reorder(state, milk_produced)) %>%
  ggplot() +
  geom_point(
    aes(x = milk_produced, y = state),
    size = 2.5, color = 'steelblue') +
  theme_minimal_vgrid() +
  labs(
      x = 'Milk produced (billions lbs)',
      y = 'State',
      title = "Top 10 milk producing states in 2017"
  )
```

## Your turn - practice plotting proportions

Using the `wildlife_impacts` data, create plots that shows the proportion of incidents that occur at each different time of day.

For this exercise, you can remove `NA` values.

Try to create the following plots:

- Stacked bars
- Dodged bars
- Pie chart
- Waffle chart

```{r}
wildlife_summary <- wildlife_impacts %>%
  filter(!is.na(time_of_day)) %>%
  count(time_of_day)

wildlife_summary
```

Stacked bars:

```{r wildlife_bars_stacked, fig.width=6, fig.height=3}
ggplot(wildlife_summary) +
  geom_col(
    aes(x = n, y = "", fill = reorder(time_of_day, n)),
    width = 0.7, alpha = 0.8) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.05))) +
  theme_minimal_hgrid() +
  labs(
    x = NULL,
    y = 'Number of incidents',
    fill = 'Time of day')
```

Dodged bars:

```{r wildlife_bars_dodged, fig.width=5, fig.height=3}
wildlife_summary %>%
  ggplot() +
  geom_col(
    aes(x = n, y = reorder(time_of_day, n)),
    width = 0.7, alpha = 0.8) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.05))) +
  theme_minimal_vgrid() +
  labs(
    x = 'Time of day',
    y = 'Number of incidents',
    fill = 'Time of day')
```

Pie chart:

```{r wildlife_pie, fig.width=5, fig.height=3}
wildlife_summary %>%
  arrange(desc(n)) %>%
  mutate(
    # Reorder the time_of_day factors before computing the labels
    time_of_day = fct_reorder(time_of_day, n), 
    p = 100*(n / sum(n)),
    label = paste(round(p), '%')) %>%
  ggplot() +
  geom_col(
    aes(x = "", y = n, fill = time_of_day),
    width = 0.7, alpha = 0.8) +
  coord_polar(theta = "y") +
  geom_text(
    aes(x = 1.2, y = n, label = label),
    color = "black", size = 4,
    position = position_stack(vjust = 0.5)) +
  theme_map() +
  labs(
    x = NULL,
    y = 'Number of incidents',
    fill = 'Time of day')
```

Waffle chart:

```{r wildlife_waffle, fig.height=5, fig.width=3.5}
# Divide the number of incidents by 100
wildlife_summary %>%
  mutate(n = n / 100) %>%
  # geom_waffle() ignores factor levels and arranges the squares
  # based on the way the data frame is sorted
  arrange(desc(n)) %>% 
  ggplot() +
  geom_waffle(
    aes(fill = time_of_day, values = n),
    color = "white", size = 1, n_rows = 15,
    flip = TRUE) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  theme_minimal() +
  labs(
    fill = 'Time of day',
    x = NULL,
    y = NULL,
    title = 'Wildlife impact incidents by time of day',
    subtitle = '(1 square = 100 incidents)')
```
