---
title: "Tidy Data"
subtitle: "EMSE 4572: Exploratory Data Analysis"
date: September 07, 2022
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(readxl)
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3
)
```



# Practice 1: Reshaping Data

Run the code chunk to read in the following two data files:

- `pv_cell_production.xlsx`: Data on solar photovoltaic cell production by country
- `milk_production.csv`: Data on milk production by state

```{r}
milk_production <- read_csv(here::here('data', 'milk_production.csv'))
pv_cells <- read_excel(
    here::here('data', 'pv_cell_production.xlsx'),
    sheet = 'Cell Prod by Country',
    skip = 2
) %>%
    filter(!is.na(Year)) %>%
    # Need to convert a few columns so the data types match
    mutate(
        China = as.numeric(China),
        Taiwan = as.numeric(Taiwan),
        Malaysia = as.numeric(Malaysia),
        Germany = as.numeric(Germany),
        `South Korea` = as.numeric(`South Korea`),
        Others = as.numeric(Others)
    )
```

Now modify the format of each:

- If the data are in "wide" format, convert it to "long" with `pivot_longer()`
- If the data are in "long" format, convert it to "wide" with `pivot_wider()`

```{r}
# Write code here
```



# Practice 2: Tidy Data Wrangling

Run the code chunk to read in the following two data files:

- `gapminder.csv`: Life expectancy in different countries over time
- `gdp.csv`: GDP of different countries over time

```{r}
gapminder <- read_csv(here::here('data', 'gapminder.csv'))
gdp <- read_csv(here::here('data', 'gdp.csv'))
```

Now convert the data into a tidy (long) structure, then create the following summary data frames:

- Mean life expectancy in each year.
- Mean GDP in each year.

```{r}
# Write code here
```



# Practice 3: Tidy Data Visualization

Run the code chunk to read in the following two data files:

- `lotr_words.csv`: Word counts by character in the LOTR films
- `fed_spending.csv`: Federal R&D spending by department

```{r}
lotr <- read_csv(here::here('data', 'lotr_words.csv'))
fed_spending <- read_csv(here::here('data', 'fed_spend_wide.csv'))
```

Now convert the data into a tidy (long) structure, then create the following charts

```{r}
# Write code here
```
