---
title: "Tidy Data"
subtitle: "EMSE 4575: Exploratory Data Analysis"
date: September 07, 2022
week: 2
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(readxl)
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3
)
```



# Practice 1: Reshaping Data

Run the code chunk to read in the following two data files:

- `pv_cell_production.xlsx`: Data on solar photovoltaic cell production by country
- `milk_production.csv`: Data on milk production by state

```{r}
milk_production <- read_csv(here::here('data', 'milk_production.csv'))
pv_cells <- read_excel(
    here::here('data', 'pv_cell_production.xlsx'), 
    sheet = 'Cell Prod by Country', 
    skip = 2
) %>% 
    filter(!is.na(Year)) %>% 
    # Need to convert a few columns so the data types match
    mutate(
        China = as.numeric(China),
        Taiwan = as.numeric(Taiwan),
        Malaysia = as.numeric(Malaysia),
        Germany = as.numeric(Germany),
        `South Korea` = as.numeric(`South Korea`),
        Others = as.numeric(Others)
    )
```

Now modify the format of each:

- If the data are in "wide" format, convert it to "long" with `pivot_longer()`
- If the data are in "long" format, convert it to "wide" with `pivot_wider()`

```{r}
# milk_production is in long format - convert to wide
milk_wide <- milk_production %>%
    pivot_wider(
        names_from = state,
        values_from = milk_produced)
head(milk_wide)

# pv_cells is in wide format - convert to long
pv_cells_long <- pv_cells %>%
    pivot_longer(
        cols = China:World,
        names_to = "country",
        values_to = "numCells") %>% 
    # Remove the "World" rows as it's a total (not a country)
    filter(country != "World")

# Could also do this
pv_cells_long <- pv_cells %>%
    pivot_longer(
        cols = -Year,
        names_to = "country",
        values_to = "numCells") %>% 
    # Remove the "World" rows as it's a total (not a country)
    filter(country != "World")

head(pv_cells_long)
```



# Practice 2: Tidy Data Wrangling

Run the code chunk to read in the following two data files:

- `gapminder.csv`: Life expectancy in different countries over time
- `gdp.csv`: GDP of different countries over time

```{r}
gapminder <- read_csv(here::here('data', 'gapminder.csv'))
gdp <- read_csv(here::here('data', 'gdp.csv'))
```

Now convert the data into a tidy (long) structure, then create the following summary data frames:

- Mean life expectancy in each year.
- Mean GDP in each year.

```{r}
gapminder %>% 
    pivot_longer(
        names_to = "country", 
        values_to = "life_exp", 
        cols = -year
    ) %>% 
    group_by(year) %>% 
    summarise(mean_life_exp = mean(life_exp))
```

```{r}
gdp %>% 
    pivot_longer(
        names_to = "year", 
        values_to = "gdp", 
        cols = -country
    ) %>% 
    filter(!is.na(gdp)) %>% 
    group_by(year) %>% 
    summarise(mean_gdp = mean(gdp))
```



# Practice 3: Tidy Data Visualization

Run the code chunk to read in the following two data files:

- `lotr_words.csv`: Word counts by character in the LOTR films
- `fed_spending.csv`: Federal R&D spending by department

```{r}
lotr <- read_csv(here::here('data', 'lotr_words.csv'))
fed_spending <- read_csv(here::here('data', 'fed_spend_wide.csv'))
```

Now convert the data into a tidy (long) structure, then create the following charts

```{r}
lotr %>% 
    pivot_longer(
        cols = Female:Male, 
        names_to = "gender", 
        values_to = "word_count"
    ) %>% 
    ggplot() +
    geom_col(aes(x = word_count, y = Film, fill = gender)) +
    labs(
        x     = "Number of words spoken by characters",
        y     = 'Film',
        fill  = "Gender of character",
        title = "Male characters had far more speakings\nroles in the LOTR series films"
    ) +
    theme_bw()
```

```{r}
fed_spending %>%
    pivot_longer( 
        names_to = "department",
        values_to = "rd_budget_mil",
        cols = DOD:Other
    ) %>% 
    group_by(year) %>%
    summarise(total = sum(rd_budget_mil)) %>% 
    ggplot(aes(x = year, y = total)) +
    geom_line() +
    geom_point() + 
    labs(
        x     = "Year",
        y     = 'R&D Spending ($ M)',
        title = "Total U.S. federal R&D spending over time"
    ) +
    theme_bw()
```

