---
title: "Cleaning Data"
subtitle: "EMSE 4572: Exploratory Data Analysis"
date: September 14, 2022
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# install.packages('lubridate')
# install.packages('janitor')

library(knitr)
library(tidyverse)
library(cowplot)
library(readxl)
library(lubridate)
library(janitor)
library(here)

knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.retina = 3)
dplyr.width = Inf

# Load all the data sets:
wildlife_impacts <- read_csv(here::here('data', 'wildlife_impacts.csv'))
milk_production <- read_csv(here::here('data', 'milk_production.csv'))

# Abbreviations data, from: https://www.50states.com/abbreviations.htm
state_abbs <- tibble::tribble(
                  ~state_name,              ~state_abb,
                     "Alabama",             "AL",
                      "Alaska",             "AK",
                     "Arizona",             "AZ",
                    "Arkansas",             "AR",
                  "California",             "CA",
                    "Colorado",             "CO",
                 "Connecticut",             "CT",
                    "Delaware",             "DE",
                     "Florida",             "FL",
                     "Georgia",             "GA",
                      "Hawaii",             "HI",
                       "Idaho",             "ID",
                    "Illinois",             "IL",
                     "Indiana",             "IN",
                        "Iowa",             "IA",
                      "Kansas",             "KS",
                    "Kentucky",             "KY",
                   "Louisiana",             "LA",
                       "Maine",             "ME",
                    "Maryland",             "MD",
               "Massachusetts",             "MA",
                    "Michigan",             "MI",
                   "Minnesota",             "MN",
                 "Mississippi",             "MS",
                    "Missouri",             "MO",
                     "Montana",             "MT",
                    "Nebraska",             "NE",
                      "Nevada",             "NV",
               "New Hampshire",             "NH",
                  "New Jersey",             "NJ",
                  "New Mexico",             "NM",
                    "New York",             "NY",
              "North Carolina",             "NC",
                "North Dakota",             "ND",
                        "Ohio",             "OH",
                    "Oklahoma",             "OK",
                      "Oregon",             "OR",
                "Pennsylvania",             "PA",
                "Rhode Island",             "RI",
              "South Carolina",             "SC",
                "South Dakota",             "SD",
                   "Tennessee",             "TN",
                       "Texas",             "TX",
                        "Utah",             "UT",
                     "Vermont",             "VT",
                    "Virginia",             "VA",
                  "Washington",             "WA",
               "West Virginia",             "WV",
                   "Wisconsin",             "WI",
                     "Wyoming",             "WY",
        "District of Columbia",             "DC",
            "Marshall Islands",             "MH",
         "Armed Forces Africa",             "AE",
       "Armed Forces Americas",             "AA",
         "Armed Forces Canada",             "AE",
         "Armed Forces Europe",             "AE",
    "Armed Forces Middle East",             "AE",
        "Armed Forces Pacific",             "AP"
)
```

# 1. Joins

1) Create a data frame called `state_data` by joining the data frames `states_abbs` and `milk_production` and then selecting the variables `region`, `state_name`, `state_abb`. **Hint**: Use the `distinct()` function to drop repeated rows.

```{r}
# Approach 1 - keeps 8 extra states
state_data <- state_abbs %>%
    left_join(milk_production, by = c('state_name' = 'state')) %>%
    select(region, state_name, state_abb) %>%
    distinct()

# Approach 2
state_data <- milk_production %>%
    left_join(state_abbs, by = c('state' = 'state_name')) %>%
    select(region, state_name = state, state_abb) %>%
    distinct()

dim(state_data)
head(state_data)
```

2) Create the data frame `wildlife_impacts2` by joining the `state_data` data frame to the `wildlife_impacts` data frame, adding the variables `region` and `state_name`.

```{r}
# Approach 1
wildlife_impacts2 <- state_data %>%
    right_join(wildlife_impacts, by = c('state_abb' = 'state')) %>%
    mutate(state_abb = ifelse(state_abb == 'N/A', NA, state_abb))

# Approach 2
wildlife_impacts2 <- wildlife_impacts %>%
    left_join(state_data, by = c('state' = 'state_abb')) %>%
    rename(state_abb = state) %>%
    mutate(state_abb = ifelse(state_abb == 'N/A', NA, state_abb))

glimpse(wildlife_impacts2)
```

# 2. Re-naming and re-typing variables

Read in the `hot_dog_winners.xlsx` file and adjust the variable names and types to match those shown on the screen.

**Approach 1**: This coerces the "dogs_eaten" variables to be numbers, so any number with a "*" symbol turns into a `NA` value.

```{r}
hot_dogs <- read_excel(
    here::here('data', 'hot_dog_winners.xlsx'),
    col_types = c(
        'numeric', 'text', 'numeric', 'text',
        'text', 'numeric', 'text')
    ) %>%
    clean_names() %>%
    select(
        year,
        competitor.mens   = mens,
        competitor.womens = womens,
        dogs_eaten.mens   = dogs_eaten_3,
        dogs_eaten.womens = dogs_eaten_6,
        country.mens      = country_4,
        country.womens    = country_7)

glimpse(hot_dogs)
```

**Approach 2**: To avoid creating `NA`s with the "dogs_eaten" variables, you can skip using the `col_types` argument and just convert the variable types later using `mutate()`.

```{r}
hot_dogs <- read_excel(
    here::here('data', 'hot_dog_winners.xlsx')) %>%
    clean_names() %>%
    select(
        year,
        competitor.mens   = mens,
        competitor.womens = womens,
        dogs_eaten.mens   = dogs_eaten_3,
        dogs_eaten.womens = dogs_eaten_6,
        country.mens      = country_4,
        country.womens    = country_7) %>%
    mutate(
        dogs_eaten.mens   = parse_number(dogs_eaten.mens),
        dogs_eaten.womens = parse_number(dogs_eaten.womens),
        year              = as.numeric(year))

glimpse(hot_dogs)
```

# 3. Recoding variables

1) Use `case_when()` to modify the `phase_of_flt` variable in the `wildlife_impacts` data:

- The values `'approach'`, `'arrival'`, `'descent'`, and `'landing roll'` should be merged into a single value called `'arrival'`.
- The values `'climb'`, `'departure'`,  and `'take-off run'` should be merged into a single value called `'departure'`.
- All other values should be called `'other'`.

```{r}
arrival_names <- c('approach', 'arrival', 'descent', 'landing roll')
departure_names <- c('climb', 'departure', 'take-off run')

wildlife_impacts <- wildlife_impacts %>%
    mutate(
        phase_of_flt = str_to_lower(phase_of_flt),
        phase_of_flt = case_when(
            phase_of_flt %in% arrival_names ~ 'arrival',
            phase_of_flt %in% departure_names ~ 'departure',
            TRUE ~ 'other'
        )
    )
```

2) Use the **lubridate** package to create a new variable, `weekday_name`, from the `incident_date` variable in the `wildlife_impacts` data.

```{r}
wildlife_impacts <- wildlife_impacts %>%
    mutate(weekday_name = wday(incident_date, label = TRUE))
```

3) Use `weekday_name` and `phase_of_flt` to make this plot of "arrival" and "departure" impacts from **Feb. 2016**.

```{r, echo=FALSE, fig.width=9, fig.height=4.5, fig.align='center'}
wildlife_impacts %>%
    filter(incident_year == 2016,
           incident_month == 3) %>%
    mutate(phase_of_flt = str_to_title(phase_of_flt)) %>%
    count(weekday_name, phase_of_flt) %>%
    ggplot() +
    geom_col(aes(x = weekday_name, y = n), width = 0.8) +
    facet_wrap(~phase_of_flt, nrow = 1) +
    theme_minimal_hgrid() +
    labs(x = 'Day of the week',
         y = 'Number of incidents',
         title = 'Impacts by day of the week & phase of flight in March, 2016')
```
