---
title: "Tidy Data"
params:
  number: 1
  purpose: 'The purpose of this assignment is to introduce yourself to the "tidy data" concept and to practice modifying data frames between long and wide formats in R.'
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
source(here::here("R", "settings.R"))
```

```{r child = here::here("fragments", "hw.qmd")}
```

# 1. Software

If you haven't yet, go to the [Course Software](../help/course-software.html) page and install all the software we'll need for this course. You'll need these tools for this assignment.

# 2. Getting Organized

Download and edit [this template](`r url_template`) when working through this assignment.

# 3. Readings & Exercises

Unzip the template folder (make sure you unzip it!), then open the .Rproj file to open RStudio. Open the `hw1.Rmd` file, take notes, and write some example code as you go through the following readings / exercises:

1. **Getting Familiar with the Course**: Follow [Snoop's advice](https://www.youtube.com/watch?v=Tnlaokj1opA) and read the entire [Course Syllabus](../syllabus.html) (actually read the whole thing). Then review the [schedule](../schedule.html) and make sure to note important upcoming deadlines.
2. **Getting Familiar with Tidy Data**: Read [this tidyverse article](https://tidyr.tidyverse.org/articles/tidy-data.html) explaining the concept of tidy data.
3. **Practice Rehaping Messy Data**: Go to [this page](https://cfss.uchicago.edu/notes/tidy-exercise/) and try loading at least one of the example data sets shown. Then try rehaping the data into a tidy format.

# 4. Reflect 

Reflect on what you've learned while going through these readings and exercises. Is there anything that jumped out at you? Anything you found particularly interesting or confusing? Write at least a paragraph in your `hw1.Rmd` file.

# 5. Submit

To submit this assignment, create a zip file of all the files in your R project folder for this assignment. Name the zip file ``r paste0('hw', params$number, '-netID.zip')``, replacing `netID` with your netID (e.g., ``r paste0('hw', params$number, '-jph.zip')``). Then copy that zip file into the "submissions" folder in your Box folder created for this class (the Box folder is named your GW netID).
