---
title: "Maps"
params:
  number: 10
  purpose: 'The purpose of this assignment is to learn some techniques for creating a special type of chart: maps. We will look at different types of maps and how to work with different types of geospatial data.'
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
source(here::here("R", "settings.R"))
```

```{r child = here::here("fragments", "hw.qmd")}
```

# 1. Get Organized

Download and edit [this template](`r url_template`) when working through this assignment.

Then unzip the template folder (make sure you unzip it!), then open the .Rproj file to open RStudio. Open the ``r paste0('hw', params$number, '.Rmd')`` file, take notes, and write some example code as you go through the following.

# 2. Video

Watch [this video](https://www.youtube.com/watch?v=kIID5FDi2JQ) on map projections:

<iframe width="560" height="315" src="https://www.youtube.com/embed/kIID5FDi2JQ" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

# 3. Readings

This week we'll learn about maps and geospatial data. Maps are a unique type of data viz, and often times they are inappropriately or inaccurately used. These readings will help you learn best practices when visualizing geospatial data:

- Wilke: [15 - Visualizing geospatial data](https://serialmentor.com/dataviz/geospatial-data.html)
- [XKCD: Map projections](https://xkcd.com/977/)
- Healy: [7 - Draw maps](http://socviz.co/maps.html#maps)
- Drawing beautiful maps: Parts [one](https://www.r-spatial.org/r/2018/10/25/ggplot2-sf.html), [two](https://www.r-spatial.org/r/2018/10/25/ggplot2-sf-2.html), & [three](https://www.r-spatial.org/r/2018/10/25/ggplot2-sf-3.html)

# 4. Reflect

Reflect on what you've learned while going through these readings and exercises. Is there anything that jumped out at you? Anything you found particularly interesting or confusing? Write at least a paragraph in your ``r paste0('hw', params$number, '.Rmd')`` file. Here are some suggestions:

- Discuss some of the key insights or things you found interesting in the readings or recent class periods.
- Write about the messiest data you've seen.
- Connect the course content to your own work or project you're working on.

**This week, include in your reflection an image or link to a recent map you've seen (or one you thought was super interesting)**

# 5. Knit

Click the "knit" button to compile your ``r paste0('hw', params$number, '.Rmd')`` file into a html web page. Then open the ``r paste0('hw', params$number, '.html')`` file in a web browser and proofread your report. Does all of the formatting look correct?

# 6. Submit

To submit this assignment, create a zip file of all the files in your R project folder for this assignment. Name the zip file ``r paste0('hw', params$number, '-netID.zip')``, replacing `netID` with your netID (e.g., ``r paste0('hw', params$number, '-jph.zip')``). Then copy that zip file into the "submissions" folder in your Box folder created for this class.