---
title: "Communicating Results"
params:
  number: 12
  purpose: 'The purpose of this assignment is to learn some tips for effectively communicating the results of your analyses to others. We will look at strategies for giving effective presentations / speeches as well as best practices for designing slides.'
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
source(here::here("R", "settings.R"))
```

```{r child = here::here("fragments", "hw.qmd")}
```

# 1. Get Organized

Download and edit [this template](`r url_template`) when working through this assignment.

Then unzip the template folder (make sure you unzip it!), then open the .Rproj file to open RStudio. Open the ``r paste0('hw', params$number, '.Rmd')`` file, take notes, and write some example code as you go through the following.

# 2. Video

Watch this video on how to start a speech:

<iframe width="560" height="315" src="https://www.youtube.com/embed/w82a1FT5o88" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

# 3. Readings

We'll be discussing good techniques for telling a story this week. These readings will be a good primer:

- [Wilke Chapter 29 - Telling a story and making a point](https://serialmentor.com/dataviz/telling-a-story.html).
- Stephanie Evergreen's blog post ["So What?"](https://stephanieevergreen.com/so-what/)

# 4. Reflect

For your final reflection, in addition to any thoughts that emerged when going through the above content, write about a talk or presentation you saw that you found particularly impactful. Was there anything memorable about it that made the presenter's message stick? Do you remember how the speaker began or ended their talk?

# 5. Knit

Click the "knit" button to compile your ``r paste0('hw', params$number, '.Rmd')`` file into a html web page. Then open the ``r paste0('hw', params$number, '.html')`` file in a web browser and proofread your report. Does all of the formatting look correct?

# 6. Submit

To submit this assignment, create a zip file of all the files in your R project folder for this assignment. Name the zip file ``r paste0('hw', params$number, '-netID.zip')``, replacing `netID` with your netID (e.g., ``r paste0('hw', params$number, '-jph.zip')``). Then copy that zip file into the "submissions" folder in your Box folder created for this class.