---
title: "Cleaning Data"
params:
  number: 2
  purpose: 'The purpose of this assignment is learn some techniques for dealing with messy data. Most of the time, the raw data you get is in the wrong format and variables are not properly coded, so you will need to "clean" the data before starting any analysis. Other times the data will be split into two or more data sets, and you will need to "join" them together into a single data frame.'
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
source(here::here("R", "settings.R"))
```

```{r child = here::here("fragments", "hw.qmd")}
```

# 1. Get Organized

Download and edit [this template](`r url_template`) when working through this assignment.

# 2. Read & Practice

Unzip the template folder (make sure you unzip it!), then open the .Rproj file to open RStudio. Open the ``r paste0('hw', params$number, '.Rmd')`` file, take notes, and write some example code as you go through the following readings / exercises:

- [Stat 545 chapter 15 on joining two tables](https://stat545.com/join-cheatsheet.html)
- [Tips for Cleaning Messy Data in R](https://rfortherestofus.com/2019/12/how-to-clean-messy-data-in-r/)
- Alison Hill has an excellent free course on DataCamp called [Working with Data in the Tidyverse](https://learn.datacamp.com/courses/working-with-data-in-the-tidyverse). Unfortunately, the course was archived, but you can read a webpage version of the entire course [here](https://rpubs.com/cliex159/864632). The whole course is quite useful for learning some data cleaning skills. In particular, I suggest going through parts 2 ("Tame your data") and 4 ("Transform your data") to be best-prepared for dealing with messy real-world data.

# 3. Reflect

Reflect on what you've learned while going through these readings and exercises. Is there anything that jumped out at you? Anything you found particularly interesting or confusing? Write at least a paragraph in your ``r paste0('hw', params$number, '.Rmd')`` file. Here are some suggestions:

- Discuss some of the key insights or things you found interesting in the readings or recent class periods.
- Write about the messiest data you've seen.
- Connect the course content to your own work or project you're working on.

# 4. Knit

Click the "knit" button to compile your ``r paste0('hw', params$number, '.Rmd')`` file into a html web page. Then open the ``r paste0('hw', params$number, '.html')`` file in a web browser and proofread your report. Does all of the formatting look correct?

# 5. Submit

To submit this assignment, create a zip file of all the files in your R project folder for this assignment. Name the zip file ``r paste0('hw', params$number, '-netID.zip')``, replacing `netID` with your netID (e.g., ``r paste0('hw', params$number, '-jph.zip')``). Then copy that zip file into the "submissions" folder in your Box folder created for this class (the Box folder is named your GW netID).
