---
title: "Correlation"
params:
  number: 4
  purpose: 'The purpose of this assignment is to learn some methods to recognize and measure the correlation between two variables as well as how to find a "line of best fit" between them.'
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
source(here::here("R", "settings.R"))
```

```{r child = here::here("fragments", "hw.qmd")}
```

# 1. Get Organized

Download and edit [this template](`r url_template`) when working through this assignment.

Then unzip the template folder (make sure you unzip it!), then open the .Rproj file to open RStudio. Open the ``r paste0('hw', params$number, '.Rmd')`` file, take notes, and write some example code as you go through the following.

# 2. Readings / Videos

This week, we will learn some methods to recognize and measure the correlation between two variables as well as how to find a "line of best fit" between them. By the end, you will be able to:

- Recognize when to use correlation.
- Interpret the magnitude and direction of a correlation.
- Explain the influence of outliers.
- Describe the Pearson correlation and the Spearman correlation.
- Explain the term "line of best fit".
- Interpret intercept and slope coefficients.
- Define residuals.
- Describe the standard error of the estimate.
- Report the assumptions of linear regression and how to test them.
- Explain standardized regression and its connection to correlation.

## Readings on visualizing correlation

- [Wilke - 12: Visualizing associations](https://serialmentor.com/dataviz/visualizing-associations.html): An overview of ways to visualize associations between two or more variables.
- ["Beware Spurious Correlations", Harvard Business Review](https://hbr.org/2015/06/beware-spurious-correlations)

## Video on correlation measures

<center>
<iframe width="560" height="315" src="https://www.youtube.com/embed/fmK5YWWobH8" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>
</center>

## Optional Readings

The history of statistical correlation is rather ugly and routed in racism and eugenics. These readings discuss this topics in more detail:

- [Francis Galton pioneered scientific advances in many fields – but also founded the racist pseudoscience of eugenics](https://theconversation.com/francis-galton-pioneered-scientific-advances-in-many-fields-but-also-founded-the-racist-pseudoscience-of-eugenics-144465)
- [Column: The false, racist theory of eugenics once ruled science. Let’s never let that happen again](https://www.pbs.org/newshour/nation/column-the-false-racist-theory-of-eugenics-once-ruled-science-lets-never-let-that-happen-again)
- Listen to [this Radiolab episode](https://podcasts.apple.com/us/podcast/g-unfit/id152249110?i=1000444689017) on eugenics.

# 3. Exercises

You don't *have* to do these, but they can be really helpful for extra practice. This week, take a look at the DataCamp course ["Correlation and Regression in R"](https://learn.datacamp.com/courses/correlation-and-regression-in-r).

# 4. Reflect

Reflect on what you've learned while going through these readings and exercises. Is there anything that jumped out at you? Anything you found particularly interesting or confusing? Write at least a paragraph in your ``r paste0('hw', params$number, '.Rmd')`` file. Here are some suggestions:

- Discuss some of the key insights or things you found interesting in the readings or recent class periods.
- Write about the messiest data you've seen.
- Connect the course content to your own work or project you're working on.

# 5. Knit

Click the "knit" button to compile your ``r paste0('hw', params$number, '.Rmd')`` file into a html web page. Then open the ``r paste0('hw', params$number, '.html')`` file in a web browser and proofread your report. Does all of the formatting look correct?

# 6. Submit

To submit this assignment, create a zip file of all the files in your R project folder for this assignment. Name the zip file ``r paste0('hw', params$number, '-netID.zip')``, replacing `netID` with your netID (e.g., ``r paste0('hw', params$number, '-jph.zip')``). Then copy that zip file into the "submissions" folder in your Box folder created for this class.
