---
title: "Visualizing Information"
params:
  number: 5
  purpose: 'The purpose of this assignment is to learn what makes a "good" information visualization, which is _not_ entirely a subjective judgment. There are a wide variety of design principles available to help guide the creation of clear, effective information visualizations, many of which are rooted in research on human psychology.'
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
source(here::here("R", "settings.R"))
```

```{r child = here::here("fragments", "hw.qmd")}
```

# 1. Get Organized

Download and edit [this template](`r url_template`) when working through this assignment.

Then unzip the template folder (make sure you unzip it!), then open the .Rproj file to open RStudio. Open the ``r paste0('hw', params$number, '.Rmd')`` file, take notes, and write some example code as you go through the following.

# 2. Readings / Videos

> "Data visualization is part art and part science. The challenge is to get the art right without getting the science wrong and vice versa."
>
> -- Claus O. Wilke in [Fundamentals of Data Visualization](https://serialmentor.com/dataviz/introduction.html)

## Readings on visualizing information

These two chapters cover a lot of excellent examples of what makes a chart "good" and "bad"---the focus of what we'll talk about in class next week:

- [Healy 1: Look at Data](http://socviz.co/lookatdata.html#lookatdata): What makes figures good and bad
- [Wilke 1: Introduction](https://serialmentor.com/dataviz/introduction.html): Definitions of "ugly", "bad", and "wrong" figures

## Video on the psychology of data viz

Watch [this 40-minute video](https://www.youtube.com/watch?v=fSgEeI2Xpdc) titled "How humans see data", by John Rauser. He discusses how we can exploit our understanding of human psychology to design effective charts (I'll cover much of the same information in the video in class next week).

<center>
<iframe width="560" height="315" src="https://www.youtube.com/embed/fSgEeI2Xpdc" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>
</center>

# 3. Reflect

Reflect on what you've learned while going through these readings and exercises. Is there anything that jumped out at you? Anything you found particularly interesting or confusing? Write at least a paragraph in your ``r paste0('hw', params$number, '.Rmd')`` file. Here are some suggestions:

- Discuss some of the key insights or things you found interesting in the readings or recent class periods.
- Write about the messiest data you've seen.
- Connect the course content to your own work or project you're working on.

# 4. Knit

Click the "knit" button to compile your ``r paste0('hw', params$number, '.Rmd')`` file into a html web page. Then open the ``r paste0('hw', params$number, '.html')`` file in a web browser and proofread your report. Does all of the formatting look correct?

# 5. Submit

To submit this assignment, create a zip file of all the files in your R project folder for this assignment. Name the zip file ``r paste0('hw', params$number, '-netID.zip')``, replacing `netID` with your netID (e.g., ``r paste0('hw', params$number, '-jph.zip')``). Then copy that zip file into the "submissions" folder in your Box folder created for this class.