---
title: "Trends"
params:
  number: 8
  purpose: 'The purpose of this assignment is to learn some of the technical details of how to create charts that highlight **trends**, which inherently involve changes in values over time.'
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
source(here::here("R", "settings.R"))
```

```{r child = here::here("fragments", "hw.qmd")}
```

# 1. Get Organized

Download and edit [this template](`r url_template`) when working through this assignment.

Then unzip the template folder (make sure you unzip it!), then open the .Rproj file to open RStudio. Open the ``r paste0('hw', params$number, '.Rmd')`` file, take notes, and write some example code as you go through the following.

# 2. Readings

Most of the readings this week have code in them that illustrate how to create each chart type, and I encourage you to try and reproduce the examples provided in R yourself.

- Wilke: [13 - Visualizing time series](https://serialmentor.com/dataviz/time-series.html)
- Wilke: [14 - Visualizing trends](https://serialmentor.com/dataviz/visualizing-trends.html)

Some examples of visualizations we'll explore in class:

- The "bar chart race": [Changes in how couples meet over time](https://flowingdata.com/2019/03/15/shifts-in-how-couples-meet-online-takes-the-top-spot/)
- The heatmap: [How common is your birthday?](http://thedailyviz.com/2016/09/17/how-common-is-your-birthday-dailyviz/)
- [Time-Series Heatmaps](https://towardsdatascience.com/time-series-calendar-heatmaps-9f576578fcfe)

Optional:

- [The Sankey diagrams (a.k.a. flow chart)](https://www.data-to-viz.com/graph/sankey.html)
- Video: [The fallen of WWII](http://www.fallen.io/ww2/)

# 3. Video

This video by Hans Rosling is a great summary of his larger [TED talk](https://www.ted.com/talks/hans_rosling_the_best_stats_you_ve_ever_seen) on the "gapminder" project he started showing life expectancy and GDP over time. This is an excellent example of using an _animated_ chart to tell a much richer story than a static chart.

<iframe width="560" height="315" src="https://www.youtube.com/embed/jbkSRLYSojo" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

# 4. Reflect

Reflect on what you've learned while going through these readings and exercises. Is there anything that jumped out at you? Anything you found particularly interesting or confusing? Write at least a paragraph in your ``r paste0('hw', params$number, '.Rmd')`` file. Here are some suggestions:

- Discuss some of the key insights or things you found interesting in the readings or recent class periods.
- Write about the messiest data you've seen.
- Connect the course content to your own work or project you're working on.

# 5. Knit

Click the "knit" button to compile your ``r paste0('hw', params$number, '.Rmd')`` file into a html web page. Then open the ``r paste0('hw', params$number, '.html')`` file in a web browser and proofread your report. Does all of the formatting look correct?

# 6. Submit

To submit this assignment, create a zip file of all the files in your R project folder for this assignment. Name the zip file ``r paste0('hw', params$number, '-netID.zip')``, replacing `netID` with your netID (e.g., ``r paste0('hw', params$number, '-jph.zip')``). Then copy that zip file into the "submissions" folder in your Box folder created for this class.