---
title: "Polishing Charts"
params:
  number: 9
  purpose: 'The purpose of this assignment is to learn some techniques for adding "polish" to your charts. We will start choosing colors and fonts more carefully, and we will also annotate our charts with informative text.'
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
source(here::here("R", "settings.R"))
```

```{r child = here::here("fragments", "hw.qmd")}
```

# 1. Get Organized

Download and edit [this template](`r url_template`) when working through this assignment.

Then unzip the template folder (make sure you unzip it!), then open the .Rproj file to open RStudio. Open the ``r paste0('hw', params$number, '.Rmd')`` file, take notes, and write some example code as you go through the following.

# 2. Readings

This week we'll get into the weeds with techniques for polishing up your charts. This is a critically important step for making your charts easy for others to understand and effective as a central tool for communicating important results. It's also where you can be highly creative!

- How to polish ggplot charts: [R4DS Chapter 28 - Graphics for communication](https://r4ds.had.co.nz/graphics-for-communication.html)
- Wilke: [4 - Color scales](https://clauswilke.com/dataviz/color-basics.html)
- Rost: [Which color scale to use when visualizing data](https://blog.datawrapper.de/which-color-scale-to-use-in-data-vis/)

# Optional

- Wilke: [21 - Multi-panel figures](https://serialmentor.com/dataviz/multi-panel-figures.html)
- Check out the [patchwork package](https://patchwork.data-imaginist.com/articles/patchwork.html) for merging multiple charts together
- How to make a multi-panel [figure with a shared legend](https://wilkelab.org/cowplot/articles/shared_legends.html).
- Check out this [tool](https://github.com/clauswilke/colorblindr) for simulating color blindness.
- Check out these tools for customizing your Rmd file: [PIMP my Rmd](https://holtzy.github.io/Pimp-my-rmd/)

# 3. Practice polishing

Go back to a previous homework where you made a chart (e.g. [hw6](https://eda.seas.gwu.edu/2022-Fall/hw/6-amounts-proportions.html)). Using some of the principles in the readings you just read, try to make a more polished version of your figure.

# 3. Reflect

Reflect on what you've learned while going through these readings and exercises. Is there anything that jumped out at you? Anything you found particularly interesting or confusing? Write at least a paragraph in your ``r paste0('hw', params$number, '.Rmd')`` file. Here are some suggestions:

- Discuss some of the key insights or things you found interesting in the readings or recent class periods.
- Write about the messiest data you've seen.
- Connect the course content to your own work or project you're working on.

**Do you have a favorite font / color / color palette? Share it in your reflection this week!**

# 4. Knit

Click the "knit" button to compile your ``r paste0('hw', params$number, '.Rmd')`` file into a html web page. Then open the ``r paste0('hw', params$number, '.html')`` file in a web browser and proofread your report. Does all of the formatting look correct?

# 5. Submit

To submit this assignment, create a zip file of all the files in your R project folder for this assignment. Name the zip file ``r paste0('hw', params$number, '-netID.zip')``, replacing `netID` with your netID (e.g., ``r paste0('hw', params$number, '-jph.zip')``). Then copy that zip file into the "submissions" folder in your Box folder created for this class.