---
title: "Correlation"
subtitle: "EMSE 4572: Exploratory Data Analysis"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    comment = "#>",
    fig.path = "figs/", # Plot save path
    fig.width = 7.252, # Plot dimensions
    fig.height = 4, #<<
    fig.retina = 3 # Better plot resolution
)

# install.packages('HistData')
# install.packages('palmerpenguins')
# install.packages('GGally')

library(HistData)
library(GGally)
library(palmerpenguins)
library(tidyverse)
library(here)

# Set theme for all ggplots
theme_set(theme_bw(base_size = 20))

# Read in data
msleep <- read_csv(here('data', 'msleep.csv'))
```

# 1. Correlation

Using the `penguins` data frame:

1. Find the two variables with the largest correlation in absolute value (i.e. closest to -1 or 1).

2. Create a scatter plot of those two variables.

3. Add an annotation for the Pearson correlation coefficient.

```{r}
# Write code here
```

# 2. Linear models

Using the `msleep` data frame:

1. Create a scatter plot of `brainwt` versus `bodywt`.
2. Include an annotation for the Pearson correlation coefficient.
3. Include an annotation for the best fit line.

Bonus: Compare your results to a log-linear relationship by converting the x and y variables to the log of x and y, like this:

```{r, eval=FALSE}
model <- lm(log(x) ~ log(y), data = msleep)
```

You can also convert your plot to log axes by adding these layers:

```{r, eval=FALSE}
plot +
    scale_x_log10() +
    scale_y_log10()
```

Linear model:

```{r}
# Write code here
```

Log-linear model:

```{r}
# Write code here
```

