---
title: "Correlation"
subtitle: "EMSE 4572: Exploratory Data Analysis"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    comment = "#>",
    fig.path = "figs/", # Plot save path
    fig.width = 7.252, # Plot dimensions
    fig.height = 4, #<<
    fig.retina = 3 # Better plot resolution
)

# install.packages('HistData')
# install.packages('palmerpenguins')
# install.packages('GGally')

library(HistData)
library(GGally)
library(palmerpenguins)
library(tidyverse)
library(here)

# Set theme for all ggplots
theme_set(theme_bw(base_size = 20)) 

# Read in data
msleep <- read_csv(here('data', 'msleep.csv'))
```

# 1. Correlation

Using the `penguins` data frame:

1. Find the two variables with the largest correlation in absolute value (i.e. closest to -1 or 1).

2. Create a scatter plot of those two variables.

3. Add an annotation for the Pearson correlation coefficient.

```{r}
penguins %>%
    ggcorr(label = TRUE)

corr <- cor(
    penguins$flipper_length_mm, penguins$body_mass_g,
    use = "complete.obs")

corLab <- paste("r = ", round(corr, 2))

penguins %>% 
    ggplot() +
    geom_point(
        aes(
            x = flipper_length_mm, 
            y = body_mass_g), 
        alpha = 0.7
    ) +
    annotate(
        geom = 'text', 
        x = 175, y = 6000,
        label = corLab,
        hjust = 0, size = 5) +
    theme_classic(base_size = 18) +
    labs(
        x = "Flipper length (mm)",
        y = "Body mass (g)"
    )
```

# 2. Linear models

Using the `msleep` data frame:

1. Create a scatter plot of `brainwt` versus `bodywt`.
2. Include an annotation for the Pearson correlation coefficient.
3. Include an annotation for the best fit line.

Bonus: Compare your results to a log-linear relationship by converting the x and y variables to the log of x and y, like this:

```{r, eval=FALSE}
model <- lm(log(x) ~ log(y), data = msleep)
```

You can also convert your plot to log axes by adding these layers:

```{r, eval=FALSE}
plot +
    scale_x_log10() +
    scale_y_log10()
```

Linear model:

```{r}
# Make the correlation label
corr <- cor(
    msleep$bodywt, msleep$brainwt,
    method = 'pearson', use = "complete.obs")
corrLabel <- paste('r = ', round(corr, 2)) 

# Make the model label
modelLinear <- lm(brainwt ~ bodywt, data = msleep)
coefs <- round(coef(modelLinear), 3)
modelLabel <- str_c('y = ', coefs[1], ' + ', coefs[2], 'x')

# Make the chart
msleep %>% 
    ggplot(
        aes(
            x = bodywt, 
            y = brainwt
        )
    ) +
    geom_point() +
    geom_smooth(method = 'lm', se = FALSE) +
    annotate(
        geom = 'text', 
        x = 100, y = 6,
        label = corrLabel,
        hjust = 0, size = 5) +
    annotate(
        geom = 'text', 
        x = 100, y = 5.5,
        label = modelLabel,
        hjust = 0, size = 5) + 
    labs(
        x = "Body weight", 
        y = "Brain weight"
    )
```

Log-linear model:

```{r}
# Make the correlation label
corrLog <- cor(
    log(msleep$bodywt), log(msleep$brainwt),
    method = 'pearson', use = "complete.obs")
corrLabelLog <- paste('r = ', round(corrLog, 2)) 

# Make the model label
modelLog <- lm(log(brainwt) ~ log(bodywt), data = msleep)
coefs <- round(coef(modelLog), 3)
modelLabelLog <- str_c('log(y) = ', coefs[1], ' + ', coefs[2], ' log(x)')

# Make the chart
msleep %>% 
    ggplot(
        aes(
            x = bodywt, 
            y = brainwt
        )
    ) +
    geom_point() +
    geom_smooth(method = 'lm', se = FALSE) +
    annotate(
        geom = 'text', 
        x = 0.01, y = 6,
        label = corrLabelLog,
        hjust = 0, size = 5) +
    annotate(
        geom = 'text', 
        x = 0.01, y = 3,
        label = modelLabelLog,
        hjust = 0, size = 5) +
    labs(
        x = "log(body weight)", 
        y = "log(brain weight)"
    ) +
    scale_x_log10() +
    scale_y_log10()
```

