---
title: "Getting Started"
subtitle: "EMSE 4572: Exploratory Data Analysis"
date: August 31, 2022
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(readxl)
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3
)
```

```{r libraries, echo=FALSE}
# If you don't have these packages installed, run these lines to do so:
# install.packages('tidyverse')
# install.packages('here')
# install.packages('readxl')

# Load libraries
library(tidyverse)
library(here)
library(readxl)
```


# Importing Data

Write code to import the following data files from the "data" folder:

- `lotr_words.csv`
- `north_america_bear_killings.txt`
- `uspto_clean_energy_patents.xlsx`

```{r}
# Write code here
```


# Data Wrangling

## Quickly previewing a data frame

Some quick practice viewing pieces of data frames.

```{r}
# Extract rows and columns using df[row, col]
lotr[1,] # First row
lotr[,1] # First column

# Extract whole columns using df$name
lotr$Race
lotr$Film

# Get the dimensions
nrow(lotr)  # number of rows
ncol(lotr)  # number of columns
dim(lotr)   # numbers of rows, columns

# View dataset in new tab
view(lotr) # This is "read-only" so you can't corrupt the data :)

# View data types of each column
glimpse(lotr)
```

## Practice summarizing data frames

Read in a data set:

```{r}
data <- read_csv(here('data', 'data.csv'))
```

How many rows and columns are in the data frame?

```{r}
# Write code here
```

What type of data is each column? (Just look, don't need to type out the answer)

```{r}
# Write code here
```

Preview the different columns - what do you think this data is about? What might one row represent?

```{r}
# Write code here
```

How many unique airlines are in the data frame? 

```{r}
# Write code here
```

What is the shortest and longest air time for any one flight in the data frame?

```{r}
# Write code here
```

## Data wrangling practice

Read in the `data.csv` file in the "data" folder:

```{r, eval=FALSE}
data <- read_csv(here('data', 'data.csv'))
```

Now answer these questions:

Create a new data frame, `flights_fall`, that contains only flights that departed in the fall semester.

```{r}
# Write code here
```

Create a new data frame, `flights_dc`, that contains only flights that flew to DC airports (Reagan or Dulles).

```{r}
# Write code here
```

Create a new data frame, `flights_dc_carrier`, that contains only flights that flew to DC airports (Reagan or Dulles) and only the columns about the month and airline.

```{r}
# Write code here
```

How many unique airlines were flying to DC airports in July?

```{r}
# Write code here
```

Create a new variable, `speed`, in miles per hour using the `time` (minutes) and `distance` (miles) variables. 

```{r}
# Write code here
```

Which flight flew the fastest?

```{r}
# Write code here
```

Remove rows that have `NA` for `air_time` and re-arrange the resulting data frame based on the longest air time and longest flight distance.

```{r}
# Write code here
```

# Data Visualization

```{r}
# Write code here
```

```{r}
# Write code here
```

```{r}
# Write code here
```

## Extra practice 

```{r, fig.height=5, fig.width=7}
# Write code here
```

```{r, fig.height=4, fig.width=6}
# Write code here
```
