---
title: "Centrality & Variability"
subtitle: "EMSE 4572: Exploratory Data Analysis"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    comment = "#>",
    fig.path = "figs/", # Plot save path
    fig.width = 7.252, # Plot dimensions
    fig.height = 4, #<<
    fig.retina = 3 # Better plot resolution
)

theme_set(theme_bw(base_size = 20)) # Set theme for all ggplots

# Read in data
milk_production <- read_csv(here('data', 'milk_production.csv'))
lotr_words <- read_csv(here('data', 'lotr_words.csv'))
marathon <- read_csv(here('data', 'marathon.csv'))
wildlife_impacts <- read_csv(here('data', 'wildlife_impacts.csv'))
faithful <- read_csv(here('data', 'faithful.csv'))
candy_rankings <- read_csv(here('data', 'candy_rankings.csv'))
```

# Quick practice: What's the data type?

Decide here:

https://docs.google.com/presentation/d/1J8UtyEwkA5QEcQQ9LCAs4EU1gyhPY7QIoIW_T_gys6o/edit?usp=sharing

```{r}
wildlife_impacts %>%
  filter(!is.na(cost_repairs_infl_adj)) %>%
  select(incident_date, time_of_day, species, cost_repairs_infl_adj)
```

- incident_date:
- time_of_day:
- species:
- cost_repairs_infl_adj:



# Practice with summary measurements

1) Read in the following data sets:

- milk_production.csv
- lotr_words.csv

 2) For each variable in each data set, if possible, summarize its

- Centrality
- Variability

```{r}
# milk_production:


# lotr_words

```



# Practice with visual summaries

1) Read in the following data sets:

- faithful.csv
- marathon.csv

2) Summarize the following variables using an appropriate chart (bar chart, histogram, and / or boxplot):

- faithful: eruptions
- faithful: waiting
- marathon: Age
- marathon: State
- marathon: Country
- marathon: `Official Time`

```{r}
# faithful: eruptions


# faithful: waiting


# marathon: Age


# marathon: State


# marathon: Country


# marathon: `Official Time`

```



# Practice with visualizing _relationships_

1) Read in the following data sets:

- marathon.csv
- wildlife_impacts.csv

2) Visualize the _relationships_ between the following variables using an appropriate chart (bar plots, scatterplots, and / or box plots):

- marathon: Age & `Official Time`
- marathon: Country & `Official Time`
- wildlife_impacts: state & operator

```{r}
# marathon: Age & `Official Time`


# marathon: Country & `Official Time`


# wildlife_impacts: state & operator

```



# Practice doing EDA

1) Read in the candy_rankings.csv data sets

2) Preview the data, note the data types and what each variable is.

3) Visualize (at least) three _relationships_ between two variables (guided by a question) using an appropriate chart:

- Bar chart
- Scatterplot
- Boxplot

```{r}

```
