---
title: "Centrality & Variability"
subtitle: "EMSE 4572: Exploratory Data Analysis"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    comment = "#>",
    fig.path = "figs/", # Plot save path
    fig.width = 7.252, # Plot dimensions
    fig.height = 4, #<<
    fig.retina = 3 # Better plot resolution
)

theme_set(theme_bw(base_size = 20)) # Set theme for all ggplots

# Read in data
milk_production <- read_csv(here('data', 'milk_production.csv'))
lotr_words <- read_csv(here('data', 'lotr_words.csv'))
marathon <- read_csv(here('data', 'marathon.csv'))
wildlife_impacts <- read_csv(here('data', 'wildlife_impacts.csv'))
faithful <- read_csv(here('data', 'faithful.csv'))
candy_rankings <- read_csv(here('data', 'candy_rankings.csv'))
```

# Quick practice: What's the data type?

Decide here:

https://docs.google.com/presentation/d/1J8UtyEwkA5QEcQQ9LCAs4EU1gyhPY7QIoIW_T_gys6o/edit?usp=sharing

```{r}
wildlife_impacts %>%
  filter(!is.na(cost_repairs_infl_adj)) %>%
  select(incident_date, time_of_day, species, cost_repairs_infl_adj)
```

- incident_date:         Interval
- time_of_day:           Ordinal
- species:               Nominal
- cost_repairs_infl_adj: Ratio



# Practice with summary measurements

1) Read in the following data sets:

- milk_production.csv
- lotr_words.csv

 2) For each variable in each data set, if possible, summarize its

- Centrality
- Variability

```{r}
# milk_production:
milk_production %>%
    summarise(
        mean   = mean(milk_produced),
        median = median(milk_produced),
        range  = max(milk_produced) - min(milk_produced),
        sd     = sd(milk_produced),
        IQR    = IQR(milk_produced)) %>%
    # Convert to millions units so easier to compare
    mutate(mean = mean / 10^6, median = median / 10^6)

# Summarize for each region
milk_production %>%
    group_by(region) %>%
    summarise(
        mean   = mean(milk_produced),
        median = median(milk_produced),
        range  = max(milk_produced) - min(milk_produced),
        sd     = sd(milk_produced),
        IQR    = IQR(milk_produced)) %>%
    # Convert to millions units so easier to compare
    mutate(mean = mean / 10^6, median = median / 10^6)

# lotr_words
lotr_words %>%
    summarise(
        mean   = mean(Female),
        median = median(Female),
        range  = max(Female) - min(Female),
        sd     = sd(Female),
        IQR    = IQR(Female))

# You can also just use the summary() function
summary(milk_production$milk_produced)
summary(milk_production$year)
summary(milk_production)

summary(lotr_words$Female)
summary(lotr_words$Male)
```



# Practice with visual summaries

1) Read in the following data sets:

- faithful.csv
- marathon.csv

2) Summarize the following variables using an appropriate chart (bar chart, histogram, and / or boxplot):

- faithful: eruptions
- faithful: waiting
- marathon: Age
- marathon: State
- marathon: Country
- marathon: `Official Time`

```{r}
# faithful:
ggplot(faithful) +
    geom_histogram(aes(x = eruptions))

ggplot(faithful) +
    geom_boxplot(aes(y = eruptions))

ggplot(faithful) +
    geom_histogram(aes(x = waiting))

# marathon
ggplot(marathon) +
    geom_histogram(aes(x = Age))

ggplot(marathon) +
    geom_boxplot(aes(y = Age))

marathon %>%
    filter(Overall < 50) %>%
    count(State) %>%
    ggplot() +
    geom_col(aes(x = n, y = reorder(State, n)))

marathon %>%
    filter(Overall < 50) %>%
    count(Country) %>%
    ggplot() +
    geom_col(aes(x = n, y = reorder(Country, n)))

ggplot(marathon) +
    geom_histogram(aes(x = `Official Time`))

ggplot(marathon) +
    geom_boxplot(aes(y = `Official Time`))
```



# Practice with visualizing _relationships_

1) Read in the following data sets:

- marathon.csv
- wildlife_impacts.csv

2) Visualize the _relationships_ between the following variables using an appropriate chart (bar plots, scatterplots, and / or box plots):

- marathon: Age & `Official Time`
- marathon: Country & `Official Time`
- wildlife_impacts: state & operator

```{r}
ggplot(marathon) +
    geom_point(aes(x = `Official Time`, y = Age), size = 0.3)

marathon %>%
    filter(!is.na(Country)) %>%
    filter(Overall < 500) %>%
    ggplot() +
    geom_boxplot(
        aes(
            x = `Official Time`, 
            y = reorder(Country, `Official Time`)
        )
    )

wildlife_impacts %>%
    filter(state != "N/A") %>%
    count(operator, state) %>%
    ggplot() +
    geom_col(
        aes(
            x = n, 
            y = reorder(state, n), 
            fill = operator
        )
    )
```



# Practice doing EDA

1) Read in the candy_rankings.csv data sets

2) Preview the data, note the data types and what each variable is.

3) Visualize (at least) three _relationships_ between two variables (guided by a question) using an appropriate chart:

- Bar chart
- Scatterplot
- Boxplot

```{r}
# candy_rankings
candy_rankings %>%
  ggplot() +
  geom_boxplot(aes(x = fruity, y = winpercent))

candy_rankings %>%
  ggplot() +
  geom_boxplot(aes(x = chocolate, y = winpercent))

candy_rankings %>%
    ggplot() +
    geom_histogram(aes(x = winpercent, fill = chocolate), alpha = 0.3)
# Looks like people rate candy with chocolate much higher!

# Reeses at the top!
candy_rankings %>%
  ggplot() +
  geom_col(
      aes(
          x = winpercent, 
          y = reorder(competitorname, winpercent)
      )
  )

# No correlation between sugar and winning
candy_rankings %>%
  ggplot() +
  geom_point(aes(x = pricepercent, y = sugarpercent))

candy_rankings %>%
  ggplot() +
  geom_point(aes(x = winpercent, y = sugarpercent))

# Interesting...seems people aren't that sensitive to price
candy_rankings %>%
  ggplot() +
  geom_point(aes(x = winpercent, y = pricepercent))
```

