---
title: "Getting Started"
subtitle: "EMSE 4575: Exploratory Data Analysis"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(readxl)
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3
)
```

```{r libraries, echo=FALSE}
# If you don't have these packages installed, run these lines to do so:
# install.packages('tidyverse')
# install.packages('here')
# install.packages('readxl')

# Load libraries
library(tidyverse)
library(here)
library(readxl)
```


# Importing Data

Write code to import the following data files from the "data" folder:

- `lotr_words.csv`
- `north_america_bear_killings.txt`
- `uspto_clean_energy_patents.xlsx`

```{r}
lotrPath <- here('data', 'lotr_words.csv')
lotr <- read_csv(lotrPath)

bearPath <- here('data', 'north_america_bear_killings.txt')
bears <- read.table(bearPath, header = TRUE)

usptoPath <- here('data', 'uspto_clean_energy_patents.xlsx')
patents <- read_excel(usptoPath, sheet = "Figure 6-45", skip = 3)
```


# Data Wrangling

## Quickly previewing a data frame

Some quick practice viewing pieces of data frames.

```{r}
# Extract rows and columns using df[row, col]
lotr[1,] # First row
lotr[,1] # First column

# Extract whole columns using df$name
lotr$Race
lotr$Film

# Get the dimensions
nrow(lotr)  # number of rows
ncol(lotr)  # number of columns
dim(lotr)   # numbers of rows, columns

# View dataset in new tab
view(lotr) # This is "read-only" so you can't corrupt the data :)

# View data types of each column
glimpse(lotr)
```

## Practice summarizing data frames

Read in a data set:

```{r}
data <- read_csv(here('data', 'data.csv'))
```

How many rows and columns are in the data frame?

```{r}
dim(data)
```

What type of data is each column? (Just look, don't need to type out the answer)

```{r}
glimpse(data)
```

Preview the different columns - what do you think this data is about? What might one row represent?

```{r}
glimpse(data)
```

How many unique airlines are in the data frame? 

```{r}
length(unique(data$carrier))
```

What is the shortest and longest air time for any one flight in the data frame?

```{r}
max(data$air_time, na.rm = TRUE)
min(data$air_time, na.rm = TRUE)
```

## Data wrangling practice

Read in the `data.csv` file in the "data" folder:

```{r, eval=FALSE}
data <- read_csv(here('data', 'data.csv'))
```

Now answer these questions:

Create a new data frame, `flights_fall`, that contains only flights that departed in the fall semester.

```{r}
flights_fall <- data %>% 
    filter(month >= 8)
```

Create a new data frame, `flights_dc`, that contains only flights that flew to DC airports (Reagan or Dulles).

```{r}
flights_dc <- data %>% 
    filter(dest %in% c('DCA', 'IAD'))

flights_dc <- data %>% 
    filter(dest == 'DCA' | dest == 'IAD') # Same result
```

Create a new data frame, `flights_dc_carrier`, that contains only flights that flew to DC airports (Reagan or Dulles) and only the columns about the month and airline.

```{r}
flights_dc_carrier <- data %>% 
    filter(dest %in% c('DCA', 'IAD')) %>% 
    select(month, carrier)
```

How many unique airlines were flying to DC airports in July?

```{r}
flights_dc_carrier %>% 
    filter(month == 7) %>% 
    distinct(carrier)

# Alternative solution
df_july <- flights_dc_carrier %>% 
    filter(month == 7) 
length(unique(df_july$carrier))
```

Create a new variable, `speed`, in miles per hour using the `time` (minutes) and `distance` (miles) variables. 

```{r}
data <- data %>% 
    mutate(speed = distance / (air_time / 60)) %>% 
    arrange(desc(speed))
head(data)
```

Which flight flew the fastest?

```{r}
# Flight 1613
```

Remove rows that have `NA` for `air_time` and re-arrange the resulting data frame based on the longest air time and longest flight distance.

```{r}
data %>% 
    filter(!is.na(air_time)) %>% 
    arrange(desc(air_time), desc(distance))
```

# Data Visualization

```{r}
mpg %>% 
  ggplot() +
  geom_smooth(aes(x = displ, y = hwy, color = drv))
```

```{r}
mpg %>% 
  count(class, drv) %>% 
  ggplot() +
  geom_col(aes(x = drv, y = n, fill = class), width = 0.7)
```

```{r}
mpg %>% 
  ggplot(aes(x = displ, y = hwy)) +
  geom_point(aes(color = class)) + 
  geom_smooth(se = FALSE)
```

## Extra practice 

```{r, fig.height=5, fig.width=7}
lotr %>%
  ggplot() +
  geom_col(aes(x = word_count, y = film, fill = gender)) +
  labs(
    x     = "Number of words spoken by characters",
    y     = 'Film',
    fill  = "Gender of character",
    title = "Male characters had far more speakings roles in the LOTR series films"
  ) +
  theme_bw()
```

```{r, fig.height=4, fig.width=6}
mpg %>%
    mutate(manufacturer = str_to_title(manufacturer)) %>%
    group_by(manufacturer) %>%
    summarise(mean_hwy = mean(hwy)) %>%
    ggplot() +
    geom_col(
      aes(x = mean_hwy, y = reorder(manufacturer, mean_hwy)), 
      width = 0.9) +
    labs(
      x = 'Highway fuel economy (mpg)',
      y = 'Vehicle manufacturer', 
      title = 'Mean fuel economy by automaker'
    ) + 
    theme_minimal()
```
