---
title: "Comparisons"
subtitle: "EMSE 4572: Exploratory Data Analysis"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# install.packages("ggrepel")
# install.packages("ggridges")
library(ggrepel)
library(ggridges)
library(tidyverse)
library(here)
library(cowplot)
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3
)

# Read in the data
gapminder        <- read_csv(here('data', 'gapminder.csv'))
milk_production  <- read_csv(here('data', 'milk_production.csv'))
internet_regions <- read_csv(here('data', 'internet_users_region.csv'))
```

### Your turn - reference lines

Use the `milk_production.csv` data to create the following charts showing differences from the mean state milk production in 2017.

```{r}
milk_summary <- milk_production %>%
  filter(year == 2017) %>%
  mutate(
    milk_produced = milk_produced / 10^9,
    state = fct_reorder(state, milk_produced))
```

```{r milk-lollipop-mean, fig.height=8, fig.width=5.5, fig.align='center'}
milk_production %>%
  filter(year == 2017) %>%
  mutate(
    milk_produced = milk_produced / 10^9,
    state = fct_reorder(state, milk_produced)) %>%
  ggplot() +
  geom_point(
    aes(x = milk_produced, y = state),
    size = 2.5, color = 'steelblue') +
  geom_vline(
    xintercept = mean(milk_summary$milk_produced),
    color = 'red', linetype = 'dashed') +
  annotate(
    'text', x = 5, y = 'Georgia',
    color = 'red', hjust = 0,
    label = 'Mean\nProduction') +
  theme_minimal_vgrid() +
  labs(x = 'Milk produced (billions lbs)',
       y = 'State')
```

```{r milk-bars-diverging, fig.height=8, fig.width=6, fig.align='center'}
milk_summary %>%
  mutate(
    milk_produced = milk_produced - mean(milk_produced),
    barColor = ifelse(milk_produced > 0, 'above', 'below')) %>%
  ggplot() +
  geom_col(
    aes(x = milk_produced, y = state, fill = barColor),
    width = 0.7) +
  scale_fill_manual(values = c('steelblue', 'sienna')) +
  theme_minimal_vgrid() +
  theme(legend.position = 'none') +
  labs(
    x = 'Difference from mean milk produced (billions lbs)',
    y = 'State')
```

Now replicate it but only for the top 10, merging all other states into an "Other" category 

```{r milk-lollipop-mean-top10, fig.height=5, fig.width=5.5, fig.align='center'}
top10 <- milk_production %>%
  filter(year == 2017) %>%
  arrange(desc(milk_produced)) %>% 
  slice(1:10)

milk_summary <- milk_production %>%
  filter(year == 2017) %>%
  mutate(
    milk_produced = milk_produced / 10^9,
    state = fct_other(state, keep = top10$state)) %>% 
  group_by(state) %>% 
  summarise(milk_produced = sum(milk_produced)) %>% 
  ungroup() %>% 
  mutate(state = fct_reorder(state, milk_produced))

ggplot(milk_summary) +
  geom_point(
    aes(x = milk_produced, y = state),
    size = 2.5, color = 'steelblue') +
  geom_vline(
    xintercept = mean(milk_summary$milk_produced),
    color = 'red', linetype = 'dashed') +
  annotate(
    'text', x = 19, y = 'California',
    color = 'red', hjust = 1,
    label = 'Mean\nProduction') +
  theme_minimal_vgrid() +
  labs(x = 'Milk produced (billions lbs)',
       y = 'State')
```

## Your turn - comparing multiple categories

Using the `internet_regions` data frame, pick a strategy and create an improved version of this chart.

Strategies:

- Dodged bars
- Facets
- Bullet chart
- Dumbell chart
- Slope chart

```{r internet-usage-bars, fig.height=6, fig.width=10, out.width="80%"}
internet_regions_compare <- internet_regions %>%
  filter(year %in% c(2000, 2015)) %>%
  mutate(
    numUsers = numUsers / 10^6,
    year = as.factor(year))

ggplot(internet_regions_compare) +
  geom_col(
    aes(x = year, y = numUsers, fill = region),
    position = "dodge") +
  labs(y = "Millions of internet users")
```

```{r internet-usage-dodged, fig.height=6, fig.width=10, out.width="80%"}
internet_regions_compare %>%
  mutate(region = fct_reorder2(region, year, -numUsers)) %>%
  ggplot() +
  geom_col(
    aes(x = numUsers, y = region, fill = year),
    position = "dodge") +
  theme_minimal_vgrid(font_size = 18) +
  scale_x_continuous(
    limits = c(0, 1200),
    expand = expansion(mult = c(0, 0.05))) +
  scale_fill_manual(values = c('grey', 'gold')) +
  labs(x = "Millions of internet users", 
       y = "Region")
```

```{r internet-usage-facets, fig.height=4, fig.width=14, out.width="80%"}
internet_regions_compare %>%
  mutate(
    region = fct_recode(region,
      "Middle East &\nNorth Africa" = "Middle East & North Africa",
      "Latin America &\n Caribbean" = "Latin America & Caribbean",
      "Europe &\n Central Asia" = "Europe & Central Asia"),
    region = fct_reorder2(region, year, -numUsers)) %>%
  ggplot() +
  geom_col(
    aes(x = year, y = numUsers, fill = year),
    width = 0.6) +
  facet_wrap(vars(region), nrow = 1) +
  theme_minimal_hgrid(font_size = 18) +
  theme(legend.position = "none") +
  scale_y_continuous(
    limits = c(0, 1200),
    expand = expansion(mult = c(0, 0.05))) +
  scale_fill_manual(values = c('grey', 'gold')) +
  labs(y = "Millions of internet users")
```

```{r internet-usage-bullet, fig.height=6, fig.width=10, out.width="80%"}
internet_regions_compare %>%
  mutate(region = fct_reorder2(region, year, -numUsers)) %>%
  spread(year, numUsers) %>%
  ggplot() +
  geom_col(
    aes(x = `2000`, y = region, fill = "2000"),
    width = 0.7) +
  geom_col(
    aes(x = `2015`, y = region, fill = "2015"),
    width = 0.3) +
  theme_minimal_vgrid(font_size = 18) +
  scale_x_continuous(
    limits = c(0, 1200),
    expand = expansion(mult = c(0, 0.05))) +
  scale_fill_manual(values = c('grey', 'black')) +
  labs(y = "Millions of internet users",
       fill = "Year")
```

```{r internet-usage-dumbbell, fig.height=4, fig.width=6, fig.align='center'}
internet_regions_compare %>%
  mutate(region = fct_reorder2(region, year, -numUsers)) %>%
  ggplot(aes(x = numUsers, y = region)) +
  geom_line(
    aes(group = region),
    color = 'lightblue', size = 1) +
    geom_point(aes(color = year), size = 2.5) +
    scale_color_manual(values = c('lightblue', 'steelblue')) +
    scale_x_continuous(limits = c(0, 1200)) +
    theme_minimal_vgrid() +
    # Remove y axis line
    theme(axis.line.y = element_blank()) +
    labs(x = 'Millions of internet users',
         y = 'Region',
         color = 'Year',
         title = 'Number of internet users by world region',
         subtitle = "(2000 - 2015)")
```

```{r internet-usage-slope, fig.height=7, fig.width=10, fig.align='center'}
internet_regions_compare %>%
  mutate(
    region = fct_reorder2(region, year, -numUsers),
    region_asia = ifelse(region == "East Asia & Pacific", "asia", "other"),
    label = paste(region, ' (', round(numUsers), ')'),
    label_left = ifelse(year == 2000, label, NA),
    label_right = ifelse(year == 2015, label, NA)) %>%
    ggplot(aes(x = year, y = numUsers,
           group = region)) +
    geom_line(aes(color = region_asia), size = 1)  +
    # Add 2000 labels (left side)
    geom_text_repel(
      aes(label = label_left),
      hjust = 1, nudge_x = -0.05,
      direction = 'y', segment.color = 'grey') +
    # Add 2015 labels (right side)
    geom_text_repel(
      aes(label = label_right),
      hjust = 0, nudge_x = 0.05,
      direction = 'y', segment.color = 'grey') +
    # Move year labels to top
    scale_x_discrete(position = 'top') +
    # Annotate & adjust theme
    scale_color_manual(values = c("red", "black")) +
    theme_minimal_grid() +
    theme(panel.grid  = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position = 'none') +
    labs(x = NULL,
         y = 'Millions of internet users',
         title = 'Number of internet users by world region',
         subtitle = "(2000 - 2015)")
```




## Your turn - comparing distributions

Use the `gapminder.csv` data to create the following charts comparing the distribution of life expectancy across countries in continents in 2007.

```{r}
gapminder_2007 <- gapminder %>%
  filter(year == 2007) %>%
  mutate(continent = fct_reorder(continent, lifeExp))
```

```{r gapminder-densities, fig.height=4.5, fig.width=6.5, fig.align='center'}
gapminder_densities <- ggplot(gapminder_2007) +
  geom_density(
    aes(x = lifeExp, y = ..count.., fill = continent),
    alpha = 0.4) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))) +
  scale_fill_brewer(palette = 'Accent') +
  theme_minimal_hgrid() +
  labs(
    x = 'Life expectancy (years)',
    y = 'Count',
    fill = 'Continent',
    title = 'Distribution of life expectancy across\ncountries in continent in 2007')

gapminder_densities
```

```{r gapminder-densities-facet, fig.height=3, fig.width=10, fig.align='center'}
gapminder_densities + 
  facet_wrap(vars(continent), nrow = 1) + 
  theme(legend.position = "none") + 
  labs(title = 'Distribution of life expectancy across countries in continent in 2007')
```

```{r gapminder-ridges, fig.height=4.5, fig.width=6.5, fig.align='center'}
gapminder_2007 %>%
    filter(continent != 'Oceania') %>%
    ggplot() +
    geom_density_ridges(
      aes(x = lifeExp, y = continent),
      scale = 1.5, alpha = 0.7) +
    scale_y_discrete(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    coord_cartesian(clip = "off") +
    theme_ridges() +
    labs(x = 'Life expectancy (years)',
         y = 'Continent',
         title = 'Distribution of life expectancy across\ncountries in continent in 2007')
```
