---
title: "Polishing Charts"
subtitle: "EMSE 4572: Exploratory Data Analysis"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# install.packages('hrbrthemes')
# remotes::install_github("clauswilke/colorblindr")

# Note: if you are using a mac and are getting errors when
#       installing the hrbrthemes library, make sure you have
#       XQuartz installed on your computer: https://www.xquartz.org/

library(hrbrthemes)
library(knitr)
library(cowplot)
library(viridis)
library(lubridate)
library(tidyverse)
library(here)

knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3
)

# Load all the data sets
wildlife_impacts <- read_csv(here::here('data', 'wildlife_impacts.csv'))
lotr_words       <- read_csv(here::here('data', 'lotr_words.csv'))
milk_production  <- read_csv(here::here('data', 'milk_production.csv'))
federal_spending <- read_csv(here::here('data', 'federal_spending_long.csv'))

# Create some summary data frames
lotr_summary <- lotr_words %>%
    gather(key = 'gender', value = 'wordCount', Female:Male) %>%
    group_by(Race, gender) %>%
    summarise(wordCount = sum(wordCount)) %>%
    ungroup() %>%
    mutate(Race = fct_reorder(Race, desc(wordCount)))

wildlife_costs <- wildlife_impacts %>%
  rename(cost = cost_repairs_infl_adj) %>%
  filter(! is.na(cost)) %>%
  mutate(
    cost = cost / 10^6,
    incident_date = as.Date(incident_date))

arrivals <- c('approach', 'arrival', 'descent', 'landing roll')
departures <- c('climb', 'departure', 'take-off run')
wildlife_impacts_2016 <- wildlife_impacts %>%
  filter(incident_year == 2016) %>%
  mutate(
    phase_of_flt = str_to_lower(phase_of_flt),
    phase_of_flt = case_when(
      phase_of_flt %in% arrivals ~ 'arrival',
      phase_of_flt %in% departures ~ 'departure',
      TRUE ~ 'other'),
    phase_of_flt = str_to_title(phase_of_flt))

milk_top10states <- milk_production %>%
  filter(year == 2017) %>%
  arrange(desc(milk_produced)) %>%
  slice(1:10) %>%
  mutate(
    milk_produced = milk_produced / 10^9,
    state = fct_reorder(state, milk_produced))

milk_summary_dumbbell <- milk_production %>%
  filter(
    year %in% c(1970, 2017),
    state %in% milk_top10states$state) %>%
  mutate(
    # Reorder state variables to follow top 10 states
    state = fct_relevel(state, levels(milk_top10states$state)),
    # Convert year to discrete variable
    year = as.factor(year),
    # Modify the units
    milk_produced = milk_produced / 10^9)
```

# 1. Scales

Adjust the scales in the code chunk provided to match the chart on the slides.

```{r, fig.height=5, fig.width=7, fig.align='center'}
ggplot(wildlife_costs) +
  geom_point(aes(x = incident_date, y = cost)) +
  scale_x_date(
    date_breaks = '3 years',
    date_labels = "'%y") +
  scale_y_continuous(
    breaks = seq(0, 16, 4),
    labels = scales::dollar) +
  theme_minimal_grid(font_size = 16) +
  labs(
    x = 'Year',
    y = 'Cost of incident repairs ($ millions)',
    title = 'Repair costs of aircraft impacts with wildlife',
    subtitle = 'Top 4 airlines',
    caption = 'Source: https://wildlife.faa.gov/home')
```

# 2. Annotations

Use the `lotr_summary` data frame to create the following chart.

Hints:
- For the `geom_label()`, use these points:
    - `x = 0.6`
    - `y = 2100`
- For the `geom_curve()`, use these points:
    - `x = 1.2`
    - `xend = 1`
    - `y = 1300`
    - `yend = 200`

```{r lotr_bars, echo=FALSE, fig.height=6, fig.width=7.5, fig.align='center'}
label <- "Across 3 films,
only 16 words were
spoken by a
female hobbit."

ggplot(lotr_summary) +
  geom_col(
    aes(x = Race, y = wordCount, fill = gender),
    width = 0.7, color = 'white') +
  facet_wrap(vars(gender)) +
  geom_curve(
    data = data.frame(
      x = 1.2, xend = 1, y = 1300, yend = 200, gender = 'Female'),
    mapping = aes(x = x, y = y, xend = xend, yend = yend),
    color = 'grey75', size = 0.5, curvature = 0.1,
    arrow = arrow(length = unit(0.01, 'npc'), type = 'closed')) +
  geom_label(
    data = data.frame(
      x = 0.6, y = 2100, label = label, gender = 'Female'),
    mapping = aes(x = x, y = y, label = label),
    hjust = 0, lineheight = .8) +
  scale_y_continuous(
    labels = scales::comma,
    expand = expand_scale(mult = c(0, 0.05))) +
  theme_minimal_hgrid(font_size = 18) +
  theme(legend.position = 'none') +
  panel_border() +
  labs(
    x = 'Race of character',
    y = 'Number of words spoken',
    title = 'Speaking roles in the LOTR film trilogy was\nheavily biased towards male characters')
```

# 3. Colors

Make 3 different versions of this chart:

1. Change the colors to the `"RdYlBu"` ColorBrewer palette.
2. Change the colors to the `"inferno"` palette from the **viridis** library.
3. Change the colors to your custom triadic palette:

    - Use the ["eye dropper"](https://chrome.google.com/webstore/detail/eye-dropper/hmdcmlfkchdmnmnmheododdhjedfccka?hl=en) tool in Google Chrome to select a color from a website, then
    - Use your color and the [color wheel tool](https://www.sessions.edu/color-calculator/) to find a triadic color palette.

Base chart:

```{r, fig.width=7, fig.height=5, fig.align='center'}
base_plot <- ggplot(wildlife_impacts_2016) +
  geom_point(
    aes(x = speed, y = height, color = phase_of_flt),
    alpha = 0.7, size = 2) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal_grid(font_size = 15) +
  labs(x = 'Speed (mph)',
       y = 'Height (ft)',
       color = 'Phase of flight',
       title = 'Wildlife impacts in 2016')

base_plot
```

Same chart but with the `"RdYlBu"` ColorBrewer palette:

```{r, fig.width=7, fig.height=5, fig.align='center'}
base_plot +
    scale_color_brewer(palette = 'RdYlBu')
```

Same chart but with the `"inferno"` palette from the **viridis** library:

```{r, fig.width=7, fig.height=5, fig.align='center'}
base_plot +
    scale_color_viridis(discrete = TRUE, option = 'inferno')
```

Same chart but with a custom color palette:

```{r, fig.width=7, fig.height=5, fig.align='center'}
colorPalette <- c('#288ecd', '#cdb428', '#cd285c')

base_plot +
    scale_color_manual(values = colorPalette)
```

# 4. Fonts

Modify the fonts and annotations in the dumbbell chart to match the chart shown here. The main font is `'Roboto Condensed'`.

Once you've recreated the plot, try other fonts and themes, such as:

- The `'Source Sans Pro'` font.
- The `'Lato'` font.
- The `theme_ipsum()` theme from the `hrbrthemes` library.

Hint: Use `annotate()` to insert the year labels at the top:

  - 1970: `x = 9`, `y = 10.5`
  - 2017: `x = 40`, `y = 10.5`

```{r, fig.height=6, fig.width=7, fig.align='center'}
milk_summary_dumbbell %>%
  ggplot(aes(x = milk_produced, y = state)) +
  geom_line(
    aes(group = state),
    color = 'lightblue', size = 1) +
  geom_point(
    aes(color = year), size = 2.5) +
  annotate(geom = "text",
    x = 8, y = 10.5, label = '1970',
    color = 'lightblue', size = 5,
    family = 'Roboto Condensed') +
  annotate(geom = "text",
    x = 42, y = 10.5, label = '2017',
    color = 'steelblue', size = 5,
    family = 'Roboto Condensed') +
  scale_color_manual(values = c('lightblue', 'steelblue')) +
  coord_cartesian(clip = 'off') +
  theme_minimal_vgrid(font_family = 'Roboto Condensed') +
  theme(
    axis.line.y = element_blank(),
    axis.ticks = element_blank(),
    legend.position = 'none',
    plot.margin = margin(0.1, 1, 0.1, 0.1, "cm")) +
  labs(
    x = 'Milk produced (billion lbs)',
    y = 'State',
    color = 'Year',
    title = 'Top 10 milk producing states\n')
```












# Other things I showed in class

## Preview colors

```{r}
colors <- c('lightblue', 'sienna', 'forestgreen', '#eae9ea')
scales::show_col(colors)
```

## Donald Duck in Mathmagic Land

https://youtu.be/U_ZHsk0-eF0

## Complete guide to scales

https://ggplot2tor.com/scales

## App for annotating your ggplots

Install

```{r, eval=FALSE}
remotes::install_github("mattcowgill/ggannotate")
```

Example

```{r, eval=FALSE}
library(ggannotate)

plot <- ggplot(mpg) +
  geom_point(aes(x = displ, y = hwy, color = as.factor(cyl))) +
  theme_half_open()

ggannotate(plot)
```

## Color wheel tool

https://www.sessions.edu/color-calculator/

## Eye dropper tool

https://chrome.google.com/webstore/detail/eye-dropper/hmdcmlfkchdmnmnmheododdhjedfccka?hl=en

## Package for testing for color blindness

Install

```{r, eval=FALSE}
# remotes::install_github("clauswilke/colorblindr")
```

Example

```{r, eval=FALSE}
library(colorblindr)

mpg_plot <- ggplot(mtcars, aes(x = mpg, y = hp)) +
  geom_point(
    aes(color = as.factor(cyl)),
    alpha = 0.8, size = 3) +
  theme_half_open(font_size = 10) +
  scale_color_manual(values = c(
    '#a0522d', '#522da0', '#2da052'))

cvd_grid(mpg_plot)
```

## Color palettes

- **viridis**: https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html
- **ColorBrewer**: https://www.datanovia.com/en/blog/the-a-z-of-rcolorbrewer-palette/
- **Inauguration palette**: https://github.com/ciannabp/inauguration
- **PNWColors**: https://github.com/jakelawlor/PNWColors

## **rmdformats package**

https://github.com/juba/rmdformats

Install

```{r, eval=FALSE}
install.packages("rmdformats")
```

Example YAML:

```
---
title: "My document"
date: "`r Sys.Date()`"
author: John Doe
output:
  rmdformats::downcute:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---
```

## Tips & tricks for Rmd files

- **Pimp my RMD**: https://holtzy.github.io/Pimp-my-rmd/

# Annotation examples of the Federal Spending data from class

Create the summary data frame first: 

```{r}
federal_spending_summary <- federal_spending %>%
    mutate(department = fct_other(
        department, keep = 'DOD')) %>%
    group_by(department, year) %>%
    summarise(rd_budget = sum(rd_budget) / 10^3) %>%
    ungroup() %>%
    mutate(department = fct_relevel(
        department, c('Other', 'DOD')))
```

Baseline chart:

```{r fed_spending, fig.height=5, fig.width=7, fig.align='center'}
dod_spending <- ggplot(federal_spending_summary) +
  geom_area(
    aes(x = year, y = rd_budget, fill = department)) +
  scale_y_continuous(
    expand = expand_scale(mult = c(0, 0.05))) +
  scale_fill_manual(values = c('grey', 'sienna')) +
  theme_minimal_hgrid() +
  labs(
    x = NULL,
    y = 'R&D Budget ($ Billions)',
    title = 'Federal R&D spending, 1976 - 2017',
    subtitle = 'Source: AAAS Historical Trends')

dod_spending
```

Annotations with `annotate()`:

```{r fed_spending_annotate, fig.height=5, fig.width=7, fig.align='center'}
dod_spending +
    annotate(geom = 'text', x = 1995, y = 85, 
             label = 'Other', size = 6, color = 'black') + 
    annotate(geom = 'text', x = 1995, y = 25, 
             label = 'DOD', size = 6, color = 'white') + 
    theme(legend.position = 'none') 
```

Annotations with `geom_text()`:

```{r fed_spending_geom_text, fig.height=5, fig.width=7, fig.align='center'}
dod_spending +
    geom_text( 
      data = data.frame(x = 1995, y = 85, label = 'Other'), 
      aes(x = x, y = y, label = label), 
      size = 6, color = 'black') + 
    geom_text( 
      data = data.frame(x = 1995, y = 25, label = 'DOD'), 
      aes(x = x, y = y, label = label), 
      size = 6, color = 'white') + 
    theme(legend.position = 'none') 
```

Annotations with `geom_label()`:

```{r fed_spending_geom_label, fig.height=5, fig.width=7, fig.align='center'}
dod_spending +
    geom_label( 
      data = data.frame(x = 1995, y = 85, label = 'Other'), 
      aes(x = x, y = y, label = label), 
      size = 6, color = 'black') + 
    geom_label( 
      data = data.frame(x = 1995, y = 25, label = 'DOD'), 
      aes(x = x, y = y, label = label), 
      size = 6, color = 'black') + 
    theme(legend.position = 'none') 
```