---
title: "title"
author: "author"
output:
  html_document:
    toc: no
    toc_float: no
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.retina = 3,
  fig.path = "figs/"
)
```

Last week we saw how to hide code chunks using the `code_folding: hide` setting in the YAML. Another option is to hide all of the code chunks by default by setting `echo = FALSE` in the `setup` code chunk and then show all of the code in an appendix. Fortunately, this can be automatically added with a single chunk at the end of the document! For example:

Let's say I had a code chunk like this:

```{r}
# View the first few rows of the data
head(mtcars)
```

and then I made a chart like this:

```{r scatterplot, fig.width=7, fig.height=4}
# Make a scatterplot
library(ggplot2)

ggplot(mtcars) + 
    geom_point(aes(x = mpg, y = hp))
```

I could then insert an appendix of all the code chunks in this document like this:

# Appendix

Below is all of the code used to generate this report

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
