---
title: "Interactive Charts"
subtitle: "EMSE 4572: Exploratory Data Analysis"
output: html_document
---

```{r install, echo=FALSE, message=FALSE, warning=FALSE}
# These packages are needed
# Use these commands to install them if you have not already
# install.packages('plotly')
# install.packages('htmlwidgets')
# install.packages('htmltools')
# install.packages('leaflet')
# install.packages('DT')
# install.packages('reactable')
# install.packages('reactablefmtr')
# install.packages('sparkline')

# Install these for maps (in case you didn't last week)
# install.packages('maps')
# install.packages('sf')
# install.packages('rgeos')
# install.packages('devtools')
# install.packages('rnaturalearth')
# devtools::install_github("ropensci/rnaturalearthhires")
# devtools::install_github("ropensci/rnaturalearthdata")
```

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3
)

library(tidyverse)
library(here)
library(knitr)
library(viridis)
library(cowplot)
library(rnaturalearth)
library(rnaturalearthhires)
library(rnaturalearthdata)
library(plotly)
library(htmlwidgets)
library(htmltools)
library(leaflet)
library(DT)
library(reactable)
library(reactablefmtr)
library(sparkline)

# Read in data sets
internet_regions <- read_csv(here('data', 'internet_users_region.csv'))
lotr_words <- read_csv(here::here('data', 'lotr_words.csv'))
gapminder <- read_csv(here::here("data", "gapminder.csv"))
internet_users <- read_csv(here::here('data', 'internet_users_country.csv'))
```

# 1. Interactive Charts

1. Open your reflection from this past week
2. Take turns sharing your interactive chart
3. With a classmate, go back to a chart we made in a previous class and make it interactive using either `ggplotly()` or `plot_ly()`




# 2. Interactive Tables

Use `reactable()` to make the following interactive table

Read [this example](https://kcuilla.github.io/reactablefmtr/articles/embed_img.html) and use the `gapminder_flags` data frame

```{r}
gapminder_flags <- gapminder %>%
  mutate(flag = paste(
    "https://flagshub.com/images/flag-of-",
    str_to_lower(country), ".png", sep = "")) %>%
  select(flag, country, everything())
```

```{r gapminder_reactable_flags, echo=FALSE}
gapminder_flags %>%
```


# 3. Interactive maps

Create the `world_internet_2015` data frame:

```{r}
world_internet_2015 <- ne_countries(
  scale = "medium", returnclass = "sf") %>%
  select(code = iso_a3) %>%
  left_join(internet_users, by = "code") %>%
  filter(year == 2015)
```

Create the color palette:

```{r}
pal <-
```

Create the labels:

```{r}
world_internet_2015 <- world_internet_2015 %>%
```

Create the leaflet map:

```{r internet_leaflet_map, fig.width=8, fig.height=6, fig.align='center'}
leaflet(data = world_internet_2015)
```
