---
title: "Interactive Charts"
subtitle: "EMSE 4572: Exploratory Data Analysis"
output: html_document
---

```{r install, echo=FALSE, message=FALSE, warning=FALSE}
# These packages are needed
# Use these commands to install them if you have not already
# install.packages('plotly')
# install.packages('htmlwidgets')
# install.packages('htmltools')
# install.packages('leaflet')
# install.packages('DT')
# install.packages('reactable')
# install.packages('reactablefmtr')
# install.packages('sparkline')

# Install these for maps (in case you didn't last week)
# install.packages('maps')
# install.packages('sf')
# install.packages('rgeos')
# install.packages('devtools')
# install.packages('rnaturalearth')
# devtools::install_github("ropensci/rnaturalearthhires")
# devtools::install_github("ropensci/rnaturalearthdata")
```

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3
)

library(tidyverse)
library(here)
library(knitr)
library(viridis)
library(cowplot)
library(rnaturalearth)
library(rnaturalearthhires)
library(rnaturalearthdata)
library(plotly)
library(htmlwidgets)
library(htmltools)
library(leaflet)
library(DT)
library(reactable)
library(reactablefmtr)
library(sparkline)

# Read in data sets
internet_regions <- read_csv(here('data', 'internet_users_region.csv'))
lotr_words <- read_csv(here::here('data', 'lotr_words.csv'))
gapminder <- read_csv(here::here("data", "gapminder.csv"))
internet_users <- read_csv(here::here('data', 'internet_users_country.csv'))
```

# 1. Interactive Charts

Here are two examples of interactive charts using `ggplotly()`

Make the static internet usage dumbbell chart:

```{r internet-usage-dumbbell, fig.height=4, fig.width=6, fig.align='center'}
internet_regions_compare <- internet_regions %>%
  filter(year %in% c(2000, 2015)) %>%
  mutate(
    numUsers = numUsers / 10^6,
    year = as.factor(year))

dumbbell_chart <- internet_regions_compare %>%
  mutate(region = fct_reorder2(region, year, -numUsers)) %>%
  ggplot(aes(x = numUsers, y = region)) +
  geom_line(
    aes(group = region),
    color = 'lightblue', size = 1) +
    geom_point(aes(color = year), size = 2.5) +
    scale_color_manual(values = c('lightblue', 'steelblue')) +
    scale_x_continuous(limits = c(0, 1200)) +
    theme_minimal_vgrid() +
    # Remove y axis line
    theme(axis.line.y = element_blank()) +
    labs(x = 'Millions of internet users',
         y = 'Region',
         color = 'Year',
         title = 'Number of internet users by world region',
         subtitle = "(2000 - 2015)")

dumbbell_chart
```

Make the interactive version

```{r}
ggplotly(dumbbell_chart)
```

Make the static chart of the LOTR word counts:

```{r lotr_bars, echo=FALSE, fig.height=4.5, fig.width=6, fig.align='center'}
label <- "Across 3 films,
only 16 words were
spoken by a
female hobbit."

lotr_bars <- lotr_words %>%
    gather(key = 'gender', value = 'wordCount', Female:Male) %>%
    group_by(Race, gender) %>%
    summarise(wordCount = sum(wordCount)) %>%
    ungroup() %>%
    mutate(Race = fct_reorder(Race, desc(wordCount))) %>%
  ggplot() +
  geom_col(
    aes(x = Race, y = wordCount, fill = gender),
    width = 0.7, color = 'white') +
  facet_wrap(vars(gender)) +
  geom_curve(
    data = data.frame(
      x = 1.2, xend = 1, y = 1300, yend = 200, gender = 'Female'),
    mapping = aes(x = x, y = y, xend = xend, yend = yend),
    color = 'grey75', size = 0.5, curvature = 0.1,
    arrow = arrow(length = unit(0.01, 'npc'), type = 'closed')) +
  geom_label(
    data = data.frame(
      x = 0.6, y = 2100, label = label, gender = 'Female'),
    mapping = aes(x = x, y = y, label = label),
    hjust = 0, lineheight = .8) +
  scale_y_continuous(
    labels = scales::comma,
    expand = expand_scale(mult = c(0, 0.05))) +
  theme_minimal_hgrid(font_size = 10) +
  theme(legend.position = 'none') +
  panel_border() +
  labs(
    x = 'Race of character',
    y = 'Number of words spoken',
    title = 'Speaking roles in the LOTR film trilogy was heavily biased towards male characters\n\n')

lotr_bars
```

Make the interactive chart:

```{r}
ggplotly(
  lotr_bars,
  tooltip = c("Race", "wordCount"),
)
```





# 2. Interactive Tables

Use `reactable()` to make the following interactive table

Read [this example](https://kcuilla.github.io/reactablefmtr/articles/embed_img.html) and use the `gapminder_flags` data frame

```{r}
gapminder_flags <- gapminder %>%
  mutate(flag = paste(
    "https://flagshub.com/images/flag-of-",
    str_to_lower(country), ".png", sep = "")) %>%
  select(flag, country, everything())
```

```{r gapminder_reactable_flags, echo=FALSE}
gapminder_flags %>%
  reactable(
    searchable = TRUE,
    highlight = TRUE,
    filterable = TRUE,
    defaultPageSize = 5,
    showPageSizeOptions = TRUE,
    pageSizeOptions = c(5, 10, 15),
    columns = list(flag = colDef(cell = embed_img(
      width = "30", height = "20")))
  )
```



# 3. Interactive maps

Create the `world_internet_2015` data frame:

```{r}
world_internet_2015 <- ne_countries(
  scale = "medium", returnclass = "sf") %>%
  select(code = iso_a3) %>%
  left_join(internet_users, by = "code") %>%
  filter(year == 2015)
```

Create the color palette:

```{r}
pal <- colorBin(
  palette = "viridis",
  domain = world_internet_2015$percent)
```

Here's an alternative, continuous color palette:

```{r}
pal <- colorNumeric(
  palette = "viridis",
  domain = world_internet_2015$percent)
```

Create the labels:

```{r}
world_internet_2015 <- world_internet_2015 %>%
  mutate(label = paste(country, ": ", round(percent, 2), "%", sep = ""))
```

Create the leaflet map:

```{r internet_leaflet_map, fig.width=8, fig.height=6, fig.align='center'}
leaflet(data = world_internet_2015) %>%
  addPolygons(
    fillColor = ~pal(percent),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlight = highlightOptions(
      weight = 5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE),
    label = world_internet_2015$label) %>%
  addLegend(
    pal = pal, values = ~percent, opacity = 0.7,
    title = "% Pop w/internet access",
    position = "bottomright")
```
