---
title: "Maps"
subtitle: "EMSE 4572: Exploratory Data Analysis"
output: html_document
---

Load libraries and settings
```{r setup, message=FALSE, warning=FALSE}
# install.packages('maps')
# install.packages('sf')
# install.packages('rgeos')
# install.packages('rnaturalearth')
# devtools::install_github("ropensci/rnaturalearthhires")
# devtools::install_github("ropensci/rnaturalearthdata")

library(knitr)
library(cowplot)
library(viridis)
library(tidyverse)
library(here)
library(janitor)
library(maps)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)

options(
    htmltools.dir.version = FALSE,
    knitr.table.format = "html",
    knitr.kable.NA = '',
    tibble.width = Inf)
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.retina = 3)
dplyr.width = Inf

# Load all the data sets:
milk_production <- read_csv(here::here('data', 'milk_production.csv'))
us_coffee_shops <- read_csv(here::here('data', 'us_coffee_shops.csv'))
internet_users  <- read_csv(here::here('data', 'internet_users_country.csv'))
# Load state_abbs data frame, containing state abbreviations
source(here::here('data', 'state_abbs.R'))
```

# 1. Maps

Use the **rnaturalearth** library to extract and plot the shape files for China and Africa.

China:

```{r china, fig.width = 6, fig.height = 3.7}
# Get the shape file


# Make the plot

```

Africa:

```{r africa, fig.width = 6, fig.height = 3.7}
# Get the shape file


# Make the plot

```

# 2. Adding data to maps

Create the map of squirrels in NYC's central park using the following data:

- The `CentralPark.shp` file in the `data/central_park` folder.
- The `nyc_squirrels.csv` file in the `data` folder.

Read in the shape file and squirrels data:

```{r}
central_park <- st_read(here::here(
    'data', 'central_park', 'CentralPark.shp'))
squirrels <- read_csv(here::here('data', 'nyc_squirrels.csv'))
```

Make the plot:

```{r nyc_squirrels, fig.width = 10, fig.height = 4}
# Write code to make the map
```

# 3. Projections

Use the `internet_users_country.csv` data and the `world` data frame from the **rnaturalearth** library to create the map of internet access by country in 2015.

Hints:

- The `iso_a3` variable in the `worlds` data frame corresponds with the `code` variable in the `internet_users_country.csv` data frame (use this for joining).
- Use `scale_fill_gradient()` to fill the color:

```{r, eval=FALSE}
scale_fill_gradient(
    low = "#e7e1ef",
    high = "#dd1c77",
    na.value = "grey70",
    limits = c(0, 100))
```

Filter out the internet data to only include 2015 data:

```{r}
internet_users_2015 <- internet_users %>%
    filter(year == 2015)
```

Load the shape file of world countries:

```{r}
world <- ne_countries(
    scale = "medium",
    returnclass = "sf")
```

Join the internet user data to the world shape file:

```{r}
# Write code to joint the two data frames
```

Make the baseline map (with no specific projection)

```{r, fig.width = 6, fig.height = 3.7}
# Write code to make the map
```

Re-make the map using a Robinson projection

```{r, fig.width = 6, fig.height = 3.7}
# Write code to make the map with a Robinson projection
```

Re-make the map using a Mercator projection

```{r, fig.width = 6, fig.height = 3.7}
# Write code to make the map with a Mercator projection
```

# 4. Extra practice

Use the `us_states_cont` data frame and the `state_abbs` data frame to create a labeled map of the U.S.

Load the shape file of the continental US states & set the CRS to an Albers projection:

```{r}
us_states_cont <- ne_states(
    country = 'united states of america',
    returnclass = 'sf') %>%
    filter(! name %in% c('Alaska', 'Hawaii')) %>%
    st_transform(crs = "ESRI:102003")
```

Compute the centroids (the center point) of each state shape:

```{r}
centroids <- as.data.frame(
    st_coordinates(st_centroid(us_states_cont)))
names(centroids) <- c('label_x', 'label_y')
us_states_cont <- bind_cols(us_states_cont, centroids)
```

Join the `state_abbs` data frame to the `us_states_cont` data frame and make the plot:

```{r, fig.width = 9, fig.height = 5.5}
# Write code to make the map
```
