---
title: "Trends"
subtitle: "EMSE 4572: Exploratory Data Analysis"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# # Install new packages (only do this once!)
# install.packages('viridis')
# install.packages('gganimate')
# install.packages('magick')

library(knitr)
library(cowplot)
library(tidyverse)
library(viridis)
library(ggrepel)
library(gganimate)
library(magick)
library(here)

knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3
)
dplyr.width = Inf

# Load all the data sets:
global_temps <- read_csv(here::here('data', 'nasa_global_temps.csv'))
us_covid     <- read_csv(here::here('data', 'us_covid.csv'))
```

# 1. Visualizing single variables over time

Use the `global_temps` data frame to explore ways to visualize the change in average global temperatures.

Consider using:

- points
- lines
- areas
- smoothed lines

## Points

```{r global-temps-points, fig.height=5, fig.width=6}
ggplot(global_temps, aes(x = year, y = meanTemp)) +
    geom_point(color = 'steelblue', size = 1.5) +
    # Add a reference line at 0
    geom_hline(yintercept = 0) +
    theme_half_open(font_size = 18) +
    labs(x = 'Year',
         y = 'Temperature anomaly (C)')
```

## Line

```{r global-temps-line, fig.height=5, fig.width=6}
ggplot(global_temps, aes(x = year, y = meanTemp)) +
    geom_line(color = 'steelblue', size = 1) +
    # Add a reference line at 0
    geom_hline(yintercept = 0) +
    theme_half_open(font_size = 18) +
    labs(x = 'Year',
         y = 'Temperature anomaly (C)')
```

## Smoothed line

```{r global-temps-line-smoothed, fig.height=5, fig.width=6}
ggplot(global_temps, aes(x = year, y = meanTemp)) +
    geom_point(color = 'grey', size = 1.5, alpha = 0.6) +
    geom_smooth(color = 'steelblue', size = 1, se = FALSE) +
    # Add a reference line at 0
    geom_hline(yintercept = 0) +
    theme_half_open(font_size = 18) +
    labs(x = 'Year',
         y = 'Temperature anomaly (C)')
```

## Area

```{r global-temps-area, fig.height=5, fig.width=6}
ggplot(global_temps, aes(x = year, y = meanTemp)) +
    geom_area(fill = 'steelblue', alpha = 0.5) +
    geom_line(color = 'steelblue', size = 1) +
    theme_half_open(font_size = 18) +
    labs(x = 'Year',
         y = 'Temperature anomaly (C)')
```

# 2. Animation

Use the `global_temps` data frame to animate one of the charts we did in the first exercise on changes in average global temperatures.

## Points

```{r global-temps-points-anim, fig.height=5, fig.width=6}
plot <- ggplot(global_temps, aes(x = year, y = meanTemp)) +
    geom_point(color = 'steelblue', size = 1.5) +
    # Add a reference line at 0
    geom_hline(yintercept = 0) +
    theme_half_open(font_size = 16) +
    labs(x = 'Year',
         y = 'Temperature anomaly (C)')

plot_anim <- plot +
    transition_reveal(year) +
    # Leave a trail of the points
    shadow_trail(0.01)

# Note that I've specified the renderer here to use the magick library,
# which should work better on most computers
animate(plot_anim, end_pause = 10, duration = 8,
        width = 600, height = 500, res = 75, 
        renderer = magick_renderer())
```

## Line

```{r global-temps-line-anim, fig.height=5, fig.width=6}
plot <- ggplot(global_temps, aes(x = year, y = meanTemp)) +
    geom_line(color = 'steelblue', size = 1) +
    # Add a reference line at 0
    geom_hline(yintercept = 0) +
    theme_half_open(font_size = 18) +
    labs(x = 'Year',
         y = 'Temperature anomaly (C)')

plot_anim <- plot +
    transition_reveal(year)

animate(plot_anim, end_pause = 10, duration = 8,
        width = 600, height = 500, res = 75,
        renderer = magick_renderer())
```

## Area

```{r global-temps-area-anim, fig.height=5, fig.width=6}
plot <- ggplot(global_temps, aes(x = year, y = meanTemp)) +
    geom_area(fill = 'steelblue', alpha = 0.5) +
    geom_line(color = 'steelblue', size = 1) +
    theme_half_open(font_size = 18) +
    labs(x = 'Year',
         y = 'Temperature anomaly (C)')

plot_anim <- plot +
    transition_reveal(year)

animate(plot_anim, end_pause = 10, duration = 8,
        width = 600, height = 500, res = 75,
        renderer = magick_renderer())
```

# 3. Visualizing mutliple variables over time

Use the `us_covid` data frame to explore ways to visualize the number of daily cases using:

1. A labeled line chart
2. A stacked area chart
3. A heat map

## Labeled line chart

```{r covid_ca_ny, echo=FALSE, fig.height=4, fig.width=12, fig.align='center'}
us_covid_ca_ny <- us_covid %>%
  mutate(cases_daily = cases_daily / 10^3) %>%
  filter(state %in% c("California", "New York"))

ggplot(us_covid_ca_ny,
       aes(x = day, y = cases_daily, color = state)) +
  geom_line(
    aes(group = state)) +
  geom_text_repel(
    data = us_covid_ca_ny %>%
      filter(day == max(day)),
    aes(label = state),
    size = 5, direction = "y", hjust = 0, nudge_x = 5,
    segment.color = NA
  ) +
  scale_x_continuous(
    breaks = seq(0, 400, 100),
    expand = expansion(add = c(0, 50))) +
  scale_color_manual(values = c("red", "blue")) +
  theme_minimal_grid(font_size = 18) +
  theme(legend.position = "none") +
  labs(
    x = "Number of days into pandemic",
    y = 'Number of daily cases (1,000s)',
    title = 'Number of daily U.S. COVID19 cases in NY & CA')
```

## Another option: facets

```{r covid_facets, echo=FALSE, fig.height=12, fig.width=15, fig.align='center'}
us_covid %>%
  mutate(
    cases_daily = cases_daily / 10^3,
    state = fct_reorder(state, -cases_daily)) %>%
  ggplot(aes(x = day, y = cases_daily)) +
  geom_line(aes(group = state)) +
  facet_wrap(vars(state)) +
  theme_minimal_hgrid(font_size = 16) +
  theme(legend.position = "none") +
  labs(
    x = "Number of days into pandemic",
    y = 'Number of daily cases (1,000s)',
    title = 'Number of daily U.S. COVID19 cases by state')
```

## Stacked area chart

```{r covid_area, fig.height=4, fig.width=12}
us_covid %>%
  mutate(
    state = fct_other(state, keep = c(
      "California", "New York")),
    state = fct_relevel(state, c(
      "Other", "California", "New York"
    ))) %>%
  group_by(day, state) %>%
  summarise(deaths_daily = sum(deaths_daily)) %>%
  ggplot(
  aes(x = day, y = deaths_daily, fill = state)) +
  geom_area(alpha = 0.5) +
  scale_fill_manual(values = c("grey", "red", "blue")) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))) +
  theme_half_open(font_size = 18) +
  labs(
    x = "Number of days into pandemic",
    y = 'Number of daily deaths',
    fill = "State",
    title = 'Number of daily U.S. COVID19 deaths')
```

```{r covid_area_labeled, fig.height=4, fig.width=12}
covid_summary <- us_covid %>%
  mutate(
    state = fct_other(state, keep = c(
      "California", "New York")),
    state = fct_relevel(state, c(
      "Other", "California", "New York"
    ))) %>%
  group_by(day, state) %>%
  summarise(deaths_daily = sum(deaths_daily)) %>%
  ungroup

ggplot(covid_summary,
       aes(x = day, y = deaths_daily)) +
  geom_area(aes(fill = state), alpha = 0.5) +
  geom_text_repel(
    data = covid_summary %>%
      filter(day == max(day)),
    aes(label = state, color = state),
    size = 6, direction = "y", hjust = 0, nudge_x = 3,
    segment.color = NA
  ) +
  scale_x_continuous(
    breaks = seq(0, 400, 100),
    expand = expansion(add = c(0, 40))) +
  scale_fill_manual(values = c("grey", "red", "blue")) +
  scale_color_manual(values = c("grey", "red", "blue")) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))) +
  theme_half_open(font_size = 18) +
  theme(legend.position = "none") +
  labs(
    x = "Number of days into pandemic",
    y = 'Number of daily deaths',
    fill = "State",
    title = 'Number of daily U.S. COVID19 deaths')
```

## Heat map

```{r covid_heatmap, fig.height=7, fig.width=10}
us_covid %>%
    mutate(state = fct_reorder(state, cases_daily)) %>%
    ggplot() +
    geom_tile(
      aes(x = day, y = state, fill = cases_daily),
      color = 'grey80') +
    scale_x_continuous(expand = c(0, 0)) +
    scale_fill_viridis(
      option = 'inferno', direction = -1,
      trans = 'sqrt') +
    guides(fill = guide_colorbar(
      title.position = 'top',
      barwidth = 20)) +
    theme_minimal() +
    theme(
      panel.grid = element_blank(),
      legend.position = 'bottom',
      text = element_text(size = 10)) +
    labs(
        x = "Number of days into pandemic",
        y = NULL,
        fill = 'Number of daily cases',
        title = 'Number of daily U.S. COVID19 cases in each state')
```
