---
title: "Trends"
subtitle: "EMSE 4572: Exploratory Data Analysis"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# # Install new packages (only do this once!)
# install.packages('viridis')
# install.packages('gganimate')
# install.packages('gifski')
# install.packages('png')
# install.packages('magick')
# install.packages('av')

library(knitr)
library(cowplot)
library(tidyverse)
library(viridis)
library(ggrepel)
library(gganimate)
library(here)

knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3
)
dplyr.width = Inf

# Load all the data sets:
global_temps <- read_csv(here::here('data', 'nasa_global_temps.csv'))
us_covid     <- read_csv(here::here('data', 'us_covid.csv'))
```

# 1. Visualizing single variables over time

Use the `global_temps` data frame to explore ways to visualize the change in average global temperatures.

Consider using:

- points
- lines
- areas
- smoothed lines

## Points

```{r global-temps-points, fig.height=5, fig.width=6}
# Write code here
```

## Line

```{r global-temps-line, fig.height=5, fig.width=6}
# Write code here
```

## Smoothed line

```{r global-temps-line-smoothed, fig.height=5, fig.width=6}
# Write code here
```

## Area

```{r global-temps-area, fig.height=5, fig.width=6}
# Write code here
```

# 2. Animation

Use the `global_temps` data frame to animate one of the charts we did in the first exercise on changes in average global temperatures.

## Points

```{r global-temps-points-anim, fig.height=5, fig.width=6}
# Write code here
```

## Line

```{r global-temps-line-anim, fig.height=5, fig.width=6}
# Write code here
```

## Area

```{r global-temps-area-anim, fig.height=5, fig.width=6}
# Write code here
```

# 3. Visualizing mutliple variables over time

Use the `us_covid` data frame to explore ways to visualize the number of daily cases using:

1. A labeled line chart
2. A stacked area chart
3. A heat map

## Labeled line chart

For this one, try showing just the lines for NY and CA.
I already made the summary data frame to do this.

```{r covid_ca_ny, echo=FALSE, fig.height=4, fig.width=12, fig.align='center'}
us_covid_ca_ny <- us_covid %>%
  mutate(cases_daily = cases_daily / 10^3) %>%
  filter(state %in% c("California", "New York"))

# Write ggplot code here
```

## Stacked area chart

For this one, try showing NY, CA, and all other states.
I already made the summary data frame to do this.

```{r covid_area, fig.height=4, fig.width=12}
us_covid %>%
  mutate(
    state = fct_other(state, keep = c(
      "California", "New York")),
    state = fct_relevel(state, c(
      "Other", "California", "New York"
    ))) %>%
  group_by(day, state) %>%
  summarise(deaths_daily = sum(deaths_daily))

# Write ggplot code here
```

## Heat map

```{r covid_heatmap, fig.height=7, fig.width=10}
# Write code here
```
